---
title: "Getting Started with inrep"
author: "inrep Development Team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: cerulean
vignette: >
  %\VignetteIndexEntry{Getting Started with inrep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

inrep runs browser-based questionnaire studies. You describe the questions, page order, and demographic fields. inrep handles the interface, page transitions, and data collection.

This vignette covers a **non-adaptive study** — every participant sees the same items in the same order. For reporting and visualization, see **Vignette 2**. For adaptive item selection, see **Vignette 3**.

We build up the study one component at a time, assuming a single language throughout. Once each piece is in place, we cover multilingual support and assemble everything into a complete bilingual study at the end.

---

## Installation

Here we load `inrep`. In most setups, this runs quietly unless an install issue appears.

```{r install}
# Install devtools first if you don't have it yet
if (!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")

# devtools::install_github("selvastics/inrep")
library(inrep)
```

---

# How a study is built

Every inrep study script assembles the same four pieces, by default in this order:

1. **Item bank** — a data frame with one row per questionnaire item.
2. **Demographics** — a named list defining the background questions. Optional.
3. **Page flow** — a list controlling what participants see on each screen and in what order.
4. **Config + launch** — `create_study_config()` sets all study-level options; `launch_study()` starts the study.

Parts 1–3 build each piece in isolation. Part 4 introduces multilingual support. Part 5 assembles everything into one complete script.

> **This might look like a lot of configuration.** LLMs (ChatGPT, Claude, etc.) are very effective at drafting and adjusting these lists — paste in a chunk and describe what you want changed. If you prefer a point-and-click starting point, [inrep-studio](https://selvastics.shinyapps.io/inrep-studio/) generates a working study script from a web form that you can run immediately. Think of both as a blank canvas: your specific needs are best addressed through iterative refinement, with or without an LLM.

---

# Part 1: The item bank

The item bank is a plain R data frame with one row per question. inrep ships two built-in banks:

- **`bfi_items`** — 30 personality items on a 5-point Likert scale.
- **`math_items`** — binary mathematics items (correct / incorrect).

## 1.1 Looking at the built-in bank

Here we see the first rows of the item bank: question text and response options. For now, we ignore the other parameter columns because they are only used in adaptive testing (Vignette 3).

```{r load_bfi}
library(inrep)
data(bfi_items)

head(bfi_items)
```

## 1.2 Required columns

Every item bank must have at least these columns:

| Column | Required | What it holds |
|------------------------|----------|----------------------------------------------|
| `Question` | Always | The text shown to participants. |
| `ResponseCategories` | Always | The response options as a quoted comma-separated string, e.g. `"1,2,3,4,5"`. |
| `id` | Recommended | A stable label used to reference this item in scoring. |
| `domain` | Recommended | Which subscale this item belongs to. Used for grouped scoring. |
| `reverse_coded` | Recommended | `TRUE` for negatively-worded items that need flipping before averaging. |
| `a`, `b`, `b1`–`b4` | Adaptive only | IRT parameter columns used for adaptive item selection and ability estimation. Not needed for fixed-form studies — the package skips these checks automatically when `adaptive = FALSE`. See Vignette 3. |

> **`ResponseCategories` — why a string?** Use `"1,2,3,4,5"` rather than `c(1,2,3,4,5)` so the labels survive CSV export and re-import unchanged. For a 7-point scale: `"1,2,3,4,5,6,7"`.

## 1.3 Building your own item bank

When we run this chunk, we get a six-item data frame. For a fixed-form study, this is enough.

```{r custom_personality}
library(inrep)

my_items <- data.frame(
  id = paste0("C", 1:6),
  Question = c(
    "I follow through on my commitments.",
    "I lose track of important deadlines.",
    "I keep my workspace tidy.",
    "I often leave tasks half-finished.",
    "I plan ahead before acting.",
    "I get distracted from my goals easily."
  ),
  ResponseCategories = "1,2,3,4,5",
  domain        = "Conscientiousness",
  reverse_coded = c(FALSE, TRUE, FALSE, TRUE, FALSE, TRUE),
  stringsAsFactors = FALSE
)
```

For a right/wrong knowledge test, use `ResponseCategories = "0,1"`. Here we get a six-row right/wrong item bank with two domains.

```{r custom_binary}
knowledge_items <- data.frame(
  id       = paste0("Q", 1:6),
  Question = c(
    "What is 15 + 27?",  "What is 7 x 9?",
    "Solve: 2x + 3 = 11", "What is sqrt(144)?",
    "What is 25% of 200?", "What is 2^3?"
  ),
  ResponseCategories = "0,1",
  domain = rep(c("Arithmetic", "Algebra"), each = 3),
  stringsAsFactors = FALSE
)
```

---

# Part 2: The page flow

`custom_page_flow` is a plain R list. Each element describes one page. **The order of the list is the order participants see.** To reorder pages, reorder the list.

## 2.1 Four page types

| `type` | What participants see |
|----------------|-------------------------------------------------------|
| `"custom"` | Free HTML — intro, consent, transition text, or debriefing. |
| `"demographics"` | A form built automatically from `demographic_configs`. |
| `"items"` | Questionnaire items drawn from the item bank by row index. |
| `"results"` | The final screen — built-in score table and/or a custom results function. |

Every page entry needs at minimum `id`, `type`, and `title`.

## 2.2 Defining each page type

Here we define the four page types as standalone objects. Nothing renders yet — they are assembled into `custom_page_flow` in Part 5.

### Welcome page (`type = "custom"`)

A `custom` page renders whatever HTML you put in `content`. Use it for the welcome screen, consent text, transitions between blocks, or a debriefing.

```{r page_intro}
intro_page <- list(
  id      = "intro",
  type    = "custom",
  title   = "Welcome",
  content = "<h2>Welcome</h2><p>This study takes about 5 minutes.</p>"
)
```

### Demographics page (`type = "demographics"`)

inrep builds the form automatically from `demographic_configs` (Part 3). The `demographics` field lists which fields appear on this page.

```{r page_demos}
demos_page <- list(
  id           = "demos",
  type         = "demographics",
  title        = "About you",
  demographics = c("age", "gender"),   # IDs from demographic_configs
  required     = TRUE
)
```

### Item block page (`type = "items"`)

Items are referenced by **row index** in the item bank. `scale_type = "likert"` uses radio buttons; `"binary"` is for 0/1 items.

```{r page_items}
items_page <- list(
  id           = "block_1",
  type         = "items",
  title        = "Questions",
  item_indices = 1:6,              # rows 1-6 of the item bank
  scale_type   = "likert",
  instructions = "Rate how well each statement describes you."
)
```

> **Robust indices.** `item_indices = 1:6` points to rows, not item IDs. If items are added or reordered, those positions shift. Compute from IDs or domain labels instead so they stay correct after any reordering:
>
> ``` r
> c_idx <- which(my_items$id %in% paste0("C", 1:6))
> e_idx <- which(bfi_items$domain == "Extraversion")
> ```

### Results page (`type = "results"`)

Shown after the last item block. The built-in version displays a score table; custom feedback is covered in Vignette 2.

```{r page_results}
results_page <- list(
  id    = "results",
  type  = "results",
  title = "Your results"
)
```

---

# Part 3: Demographics

`demographic_configs` is a named list — one entry per background question. Entry names are what you reference in page definitions. Field types: `"number"`, `"select"`, `"checkbox"`, `"slider"`, `"text"`.

When we run this chunk, we create a named list of demographic fields. It is used later in `create_study_config()`.

```{r demo_configs}
demographic_configs <- list(

  age = list(
    question = "How old are you?",
    type     = "number",
    required = TRUE
  ),

  gender = list(
    question = "What is your gender?",
    type     = "select",
    options  = c("Female" = "f", "Male" = "m",
                 "Diverse" = "d", "Prefer not to say" = "na"),
    required = TRUE
  ),

  stress = list(
    question  = "How stressed do you feel right now?",
    type      = "slider",
    min = 0, max = 100, step = 1, default = 50,
    label_min = "Not at all",  label_max = "Extremely",
    required  = TRUE
  ),

  show_results = list(
    question = "Would you like to see your personal results at the end?",
    type     = "select",
    options  = c("Yes" = "yes", "No" = "no"),
    required = TRUE
  )
)
```

> **Slider values** are stored in the demographics output, not the responses vector. Reading a slider field from `responses[30]` returns `NA` silently. Vignette 2 covers reading demographic values.

`create_study_config()` takes two related arguments: `demographics` (the field IDs to activate) and `demographic_configs` (the full definitions). Pass both:

```{r demos_wiring}
study_config <- create_study_config(
  # ... other arguments ...
  demographics        = names(demographic_configs),
  demographic_configs = demographic_configs
)
```

---

# Part 3b: Themes

Pass one theme name to `create_study_config(theme = ...)`. That is all that is needed.

**Core themes:** `"Light"`, `"Midnight"`, `"Sunset"`, `"Forest"`, `"Ocean"`, `"Berry"`, `"Professional"`, `"hildesheim"`

**Additional themes** — inspired by popular AI tool interfaces and sourced from their public design systems: `"claude"`, `"opencode"`, `"xai"`, `"mistral"`, `"notion"`, `"linear"`, `"vercel"`, `"stripe"`, `"raycast"`, `"warp"`, `"paper"`, `"sepia"`, `"dark-mode"`, `"high-contrast"`, `"colorblind-safe"`, `"dyslexia-friendly"`, `"large-text"`, and more.

```{r themes}
create_study_config(name = "Demo", theme = "Professional", adaptive = FALSE)
create_study_config(name = "Demo", theme = "hildesheim",   adaptive = FALSE)
create_study_config(name = "Demo", theme = "claude",       adaptive = FALSE)
```

---

# Part 4: Multiple languages

So far, we have assumed the study runs in a single language and have written everything — item text, page content, demographic questions — directly in that language. In inrep, it is straightforward to either swap the entire study to a different language or to let participants choose between two languages within the same session.

Two things are controlled separately:

- **UI language** — navigation buttons, validation messages, and built-in labels. Set via `language = ...` in `create_study_config()`. Supported: `"en"`, `"de"`, `"fr"`, `"es"`.
- **Content language** — your page text, item questions, demographic questions. You write these yourself in whichever language(s) you need.

## 4.1 A single non-English study — just set `language`

Write all content in the target language in the primary fields and pass `language` to match. Here we build a short French study; navigation buttons will read *Continuer* and *Soumettre*.

```{r french_study}
french_items <- data.frame(
  id       = paste0("F", 1:4),
  Question = c(
    "Je me sens souvent fatigu\u00e9(e).",
    "Je suis satisfait(e) de ma vie en g\u00e9n\u00e9ral.",
    "Je ressens souvent de l'anxi\u00e9t\u00e9.",
    "Je me sens capable de faire face aux d\u00e9fis."
  ),
  ResponseCategories = "1,2,3,4,5",
  domain = c("Fatigue", "Satisfaction", "Anxi\u00e9t\u00e9", "Efficacit\u00e9"),
  stringsAsFactors = FALSE
)

french_flow <- list(
  list(id = "intro", type = "custom",
       title   = "Bienvenue",
       content = "<h2>Bienvenue</h2><p>Cette \u00e9tude prend environ 2 minutes.</p>"),
  list(id = "items", type = "items",
       title        = "Questions",
       item_indices = 1:4,
       scale_type   = "likert",
       instructions = "Indiquez dans quelle mesure chaque affirmation vous correspond."),
  list(id = "results", type = "results",
       title = "Vos r\u00e9sultats")
)

french_config <- create_study_config(
  name             = "\u00c9tude de bien-\u00eatre",
  custom_page_flow = french_flow,
  adaptive         = FALSE,  min_items        = 1,  max_items        = 4,
  language         = "fr"   # Continuer / Soumettre throughout
)

launch_study(config = french_config, item_bank = french_items, debug_mode = TRUE)
```

## 4.2 Bilingual studies with a toggle button

When two languages must coexist in the same session — common for studies running across language groups — every participant-facing string gets a second variant with an `_en` suffix. inrep reads the `_en` field whenever English is active and falls back to the primary field otherwise.

| Primary field | English variant |
|---|---|
| `title` | `title_en` |
| `content` | `content_en` |
| `instructions` | `instructions_en` |
| `question` (demo) | `question_en` |
| `options` (demo) | `options_en` |
| slider `label_min` / `label_max` | `label_min_en` / `label_max_en` |

For item text, add a `Question_EN` column to the item bank alongside `Question`. inrep picks the right column based on the active language.

The language is switched client-side by a button on the welcome page that calls `Shiny.setInputValue("store_language_globally", "en")`. From that point on, inrep reads the `_en` fields and switches the UI labels. Place the button on the **first** page so participants set their language before the assessment starts — `store_language_globally` takes effect on the next page render.

Here is the minimal toggle pattern:

```{r lang_toggle_example}
intro_bilingual <- list(
  id      = "intro",
  type    = "custom",
  title   = "Welcome / Willkommen",
  content = '
<div>
  <button onclick="toggleLang()" id="lang-btn"
    style="background:#005a9c;color:white;padding:8px 16px;
           border:none;border-radius:4px;cursor:pointer;font-weight:bold;">
    English Version
  </button>

  <div id="de-block">
    <h2>Willkommen</h2>
    <p>Dieser Fragebogen dauert ca. 5 Minuten.</p>
  </div>

  <div id="en-block" style="display:none;">
    <h2>Welcome</h2>
    <p>This questionnaire takes about 5 minutes.</p>
  </div>
</div>
<script>
function toggleLang() {
  var de  = document.getElementById("de-block");
  var en  = document.getElementById("en-block");
  var btn = document.getElementById("lang-btn");
  if (de.style.display === "none") {
    de.style.display = "block"; en.style.display = "none";
    btn.textContent = "English Version";
    sessionStorage.setItem("current_language", "de");
    Shiny.setInputValue("store_language_globally", "de", {priority: "event"});
  } else {
    de.style.display = "none"; en.style.display = "block";
    btn.textContent = "Deutsche Version";
    sessionStorage.setItem("current_language", "en");
    Shiny.setInputValue("store_language_globally", "en", {priority: "event"});
  }
}
</script>'
)
```

A subsequent page then carries `_en` variants on every string — inrep picks the correct one based on which language is currently active:

```{r lang_subsequent_page}
items_bilingual <- list(
  id              = "block_1",
  type            = "items",
  title           = "Fragen",
  title_en        = "Questions",
  item_indices    = 1:6,
  scale_type      = "likert",
  instructions    = "Wie gut trifft jede Aussage auf Sie zu?",
  instructions_en = "How well does each statement describe you?"
)
```

---

# Part 5: A complete study

Everything from Parts 1–4 comes together here: a bilingual item bank, demographics with `_en` variants, a page flow with a language toggle on the welcome page, a theme, and a config that ties it all to the launcher. Running the chunk opens the study in your browser; the R console shows `Listening on http://127.0.0.1:XXXX` and stays busy until you stop it.

The study runs in **German by default**. The toggle button on the welcome page switches both content and UI labels to English simultaneously. All five BFI domains are covered across three item blocks using a compact 15-item bilingual bank defined inline. Indices are derived from domain labels so they stay correct after any reordering.

**Debug shortcuts** (only when `debug_mode = TRUE`, red indicator bottom-right): **Ctrl+A** fills the current page; **Ctrl+Q** fills all pages and jumps to results.

```{r complete_study}
library(inrep)

# ============================================================
# 1. Bilingual item bank — 3 items per BFI domain (15 total).
#    Question    = German (default; shown when language = "de")
#    Question_EN = English (used when participant activates the toggle)
# ============================================================
bfi_bilingual <- data.frame(
  id = c("E1","E2","E3", "A1","A2","A3", "C1","C2","C3",
         "N1","N2","N3", "O1","O2","O3"),
  Question = c(
    "Ich gehe aus mir heraus, bin gesellig.",
    "Ich bin eher ruhig und zur\u00fcckhaltend.",
    "Ich bin begeisterungsf\u00e4hig und kann andere mitrei\u00dfen.",
    "Ich schenke anderen leicht Vertrauen.",
    "Ich bin manchmal unh\u00f6flich und unfreundlich.",
    "Ich bin hilfsbereit und unegoistisch.",
    "Ich erledige Aufgaben gr\u00fcndlich.",
    "Ich bin bequem, neige zur Faulheit.",
    "Ich arbeite zuverl\u00e4ssig und gewissenhaft.",
    "Ich bin entspannt, lasse mich durch Stress nicht aus der Ruhe bringen.",
    "Ich mache mir viele Sorgen.",
    "Ich werde leicht nerv\u00f6s und unsicher.",
    "Ich bin neugierig auf viele verschiedene Dinge.",
    "Ich habe eine aktive Vorstellungskraft, bin fantasievoll.",
    "Ich sch\u00e4tze k\u00fcnstlerische und \u00e4sthetische Eindr\u00fccke."
  ),
  Question_EN = c(
    "I am outgoing and sociable.",
    "I tend to be quiet.",
    "I am full of energy.",
    "I tend to trust others.",
    "I can sometimes be rude.",
    "I am helpful and generous.",
    "I do a thorough job.",
    "I tend to be lazy.",
    "I am efficient and reliable.",
    "I am relaxed and handle stress well.",
    "I worry a lot.",
    "I can be tense.",
    "I am curious about many different things.",
    "I have an active imagination.",
    "I appreciate artistic and aesthetic experiences."
  ),
  ResponseCategories = "1,2,3,4,5",
  domain        = rep(c("Extraversion", "Agreeableness", "Conscientiousness",
                        "Neuroticism", "Openness"), each = 3),
  reverse_coded = c(FALSE, TRUE, FALSE,
                    FALSE, TRUE, FALSE,
                    FALSE, TRUE, FALSE,
                    TRUE,  FALSE, FALSE,
                    FALSE, FALSE, FALSE),
  stringsAsFactors = FALSE
)

# ============================================================
# 2. Demographics — German primary text, English _en variants
# ============================================================
demographic_configs <- list(

  age = list(
    question    = "Wie alt sind Sie?",
    question_en = "How old are you?",
    type        = "number",
    required    = FALSE
  ),

  gender = list(
    question    = "Was ist Ihr Geschlecht?",
    question_en = "What is your gender?",
    type        = "select",
    options     = c("Weiblich" = "f", "M\u00e4nnlich" = "m",
                    "Divers" = "d",   "Keine Angabe" = "na"),
    options_en  = c("Female" = "f", "Male" = "m",
                    "Diverse" = "d", "Prefer not to say" = "na"),
    required    = FALSE
  )
)

# ============================================================
# 3. Page flow — German primary text + _en variants on every page.
#    Indices derived from domain labels so they stay correct after
#    any reordering of the item bank.
# ============================================================
e_idx <- which(bfi_bilingual$domain == "Extraversion")
a_idx <- which(bfi_bilingual$domain == "Agreeableness")
c_idx <- which(bfi_bilingual$domain == "Conscientiousness")
n_idx <- which(bfi_bilingual$domain == "Neuroticism")
o_idx <- which(bfi_bilingual$domain == "Openness")

custom_page_flow <- list(

  # -- Welcome page with DE/EN language toggle ------------------------------
  list(
    id      = "intro",
    type    = "custom",
    title   = "Willkommen / Welcome",
    content = '<div>
  <button onclick="toggleLang()" id="lang-btn"
    style="background:#005a9c;color:white;padding:8px 16px;
           border:none;border-radius:4px;cursor:pointer;font-weight:bold;">
    English Version
  </button>

  <div id="de-block">
    <h2>Willkommen</h2>
    <p>In dieser Studie untersuchen wir die Big-Five-Pers&ouml;nlichkeitseigenschaften.
    Die Teilnahme ist freiwillig und anonym. Der Fragebogen dauert ca. 5 Minuten.</p>
    <p>Bitte beurteilen Sie, wie gut jede Aussage auf Sie zutrifft
    (1 = trifft gar nicht zu, 5 = trifft voll zu).</p>
  </div>

  <div id="en-block" style="display:none;">
    <h2>Welcome</h2>
    <p>This study examines the Big Five personality traits.
    Participation is voluntary and anonymous. The questionnaire takes about 5 minutes.</p>
    <p>Please rate how well each statement describes you
    (1 = not at all, 5 = very well).</p>
  </div>
</div>
<script>
function toggleLang() {
  var de  = document.getElementById("de-block");
  var en  = document.getElementById("en-block");
  var btn = document.getElementById("lang-btn");
  if (de.style.display === "none") {
    de.style.display = "block"; en.style.display = "none";
    btn.textContent = "English Version";
    sessionStorage.setItem("current_language", "de");
    Shiny.setInputValue("store_language_globally", "de", {priority: "event"});
  } else {
    de.style.display = "none"; en.style.display = "block";
    btn.textContent = "Deutsche Version";
    sessionStorage.setItem("current_language", "en");
    Shiny.setInputValue("store_language_globally", "en", {priority: "event"});
  }
}
</script>'
  ),

  # -- Background questions -------------------------------------------------
  list(
    id           = "demos",
    type         = "demographics",
    title        = "Angaben zur Person",
    title_en     = "About you",
    demographics = names(demographic_configs),
    required     = FALSE
  ),

  # -- Item block 1: Extraversion + Agreeableness ---------------------------
  list(
    id              = "block_1",
    type            = "items",
    title           = "Teil 1 von 3",
    title_en        = "Part 1 of 3",
    item_indices    = c(e_idx, a_idx),
    scale_type      = "likert",
    instructions    = "Wie gut trifft jede Aussage auf Sie zu? (1 = gar nicht, 5 = vollst\u00e4ndig)",
    instructions_en = "How well does each statement describe you? (1 = not at all, 5 = perfectly)"
  ),

  # -- Item block 2: Conscientiousness + Neuroticism ------------------------
  list(
    id              = "block_2",
    type            = "items",
    title           = "Teil 2 von 3",
    title_en        = "Part 2 of 3",
    item_indices    = c(c_idx, n_idx),
    scale_type      = "likert",
    instructions    = "Wie gut trifft jede Aussage auf Sie zu? (1 = gar nicht, 5 = vollst\u00e4ndig)",
    instructions_en = "How well does each statement describe you? (1 = not at all, 5 = perfectly)"
  ),

  # -- Item block 3: Openness -----------------------------------------------
  list(
    id              = "block_3",
    type            = "items",
    title           = "Teil 3 von 3",
    title_en        = "Part 3 of 3",
    item_indices    = o_idx,
    scale_type      = "likert",
    instructions    = "Wie gut trifft jede Aussage auf Sie zu? (1 = gar nicht, 5 = vollst\u00e4ndig)",
    instructions_en = "How well does each statement describe you? (1 = not at all, 5 = perfectly)"
  ),

  # -- Results --------------------------------------------------------------
  list(
    id         = "results",
    type       = "results",
    title      = "Vielen Dank!",
    title_en   = "Thank you!",
    content    = "<p>Ihre Antworten wurden gespeichert.</p>",
    content_en = "<p>Your responses have been saved.</p>"
  )
)

# ============================================================
# 4. Study config
# language = "de"  -> German UI labels by default (Weiter, Absenden, ...).
# The toggle button on the intro page fires
# Shiny.setInputValue("store_language_globally", "en"), after which
# inrep reads the _en fields on all subsequent pages.
# ============================================================
all_idx <- c(e_idx, a_idx, c_idx, n_idx, o_idx)

study_config <- create_study_config(
  name                = "Big Five Pers\u00f6nlichkeitsstudie", 
  custom_page_flow    = custom_page_flow,
  demographics        = names(demographic_configs),
  demographic_configs = demographic_configs,
  adaptive            = FALSE,
  min_items           = length(all_idx),
  max_items           = length(all_idx),
  language            = "de",           # German default; toggle switches to "en"
  theme               = "hildesheim",
  response_ui_type    = "radio",
  progress_style      = "bar",
  session_save        = TRUE
)

# ============================================================
# 5. Launch
# debug_mode = TRUE  -> Ctrl+Q fills all pages, jumps to results.
# Set debug_mode = FALSE before sharing the study link.
# ============================================================
launch_study(config = study_config, item_bank = bfi_bilingual, debug_mode = TRUE)
```
