---
title: "Getting Started with inrep"
author: "inrep Development Team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: cerulean
vignette: >
  %\VignetteIndexEntry{Getting Started with inrep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

inrep runs browser-based questionnaire studies. You describe the questions, page order, and demographic fields. inrep handles the interface, page transitions, and data collection.

This vignette covers a **non-adaptive study** — every participant sees the same items in the same order. For reporting and visualization, see **Vignette 2**. For adaptive item selection, see **Vignette 3**.

---

## Installation

Here we load `inrep`. In most setups, this runs quietly unless an install issue appears.

```{r install}
# Install devtools first if you don't have it yet
if (!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")

# devtools::install_github("selvastics/inrep")
library(inrep)
```

---

# How a study is built

Every inrep study script assembles the same four pieces, by default in this order:

1. **Item bank** — a data frame with one row per questionnaire item.
2. **Demographics** — a named list defining the background questions. Optional.
3. **Page flow** — a list controlling what participants see on each screen and in what order.
4. **Config + launch** — `create_study_config()` sets all study-level options; `launch_study()` starts the study.

Parts 1–3 build each piece in isolation. Part 4 assembles them into one complete script.

---

# Part 1: The item bank

The item bank is a plain R data frame with one row per question. inrep ships two built-in banks:

- **`bfi_items`** — 30 personality items on a 5-point Likert scale.
- **`math_items`** — binary mathematics items (correct / incorrect).

## 1.1 Looking at the built-in bank

Here we see the first rows of the item bank: question text and response options. For now, we ignore the other parameter columns because they are only used in adaptive testing (Vignette 3).

```{r load_bfi}
library(inrep)
data(bfi_items)

head(bfi_items)
```

## 1.2 Required columns

Every item bank must have at least these columns:

| Column | Required | What it holds |
|------------------------|----------|----------------------------------------------|
| `Question` | Always | The text shown to participants. For bilingual studies use `Question_DE` and `Question_EN`. |
| `ResponseCategories` | Always | The response options as a quoted comma-separated string, e.g. `"1,2,3,4,5"`. |
| `id` | Recommended | A stable label used to reference this item in scoring. |
| `domain` | Recommended | Which subscale this item belongs to. Used for grouped scoring. |
| `reverse_coded` | Recommended | `TRUE` for negatively-worded items that need flipping before averaging. |
| `a`, `b`, `b1`–`b4` | Adaptive only | IRT parameter columns used for adaptive item selection and ability estimation. Not needed for fixed-form studies — the package skips these checks automatically when `adaptive = FALSE`. See Vignette 3. |

> **`ResponseCategories` — why a string?** Use `"1,2,3,4,5"` rather than `c(1,2,3,4,5)` so the labels survive CSV export and re-import unchanged. For a 7-point scale: `"1,2,3,4,5,6,7"`.

## 1.3 Building your own item bank

When we run this chunk, we get a six-item data frame. For a fixed-form study, this is enough.

```{r custom_personality}
library(inrep)

my_items <- data.frame(
  id = paste0("C", 1:6),
  Question = c(
    "I follow through on my commitments.",
    "I lose track of important deadlines.",
    "I keep my workspace tidy.",
    "I often leave tasks half-finished.",
    "I plan ahead before acting.",
    "I get distracted from my goals easily."
  ),
  ResponseCategories = "1,2,3,4,5",
  domain        = "Conscientiousness",
  reverse_coded = c(FALSE, TRUE, FALSE, TRUE, FALSE, TRUE),
  stringsAsFactors = FALSE
)
```

For a right/wrong knowledge test, use `ResponseCategories = "0,1"`. The `a` and `b` columns shown below are IRT parameters — only needed if you later switch this bank to adaptive mode:

Here we get a six-row right/wrong item bank with two domains.

```{r custom_binary}
knowledge_items <- data.frame(
  id       = paste0("Q", 1:6),
  Question = c(
    "What is 15 + 27?",  "What is 7 x 9?",
    "Solve: 2x + 3 = 11", "What is sqrt(144)?",
    "What is 25% of 200?", "What is 2^3?"
  ),
  ResponseCategories = "0,1",
  a      = 1,
  b      = c(-0.5, 0.2, 0.8, -0.3, 0.1, 0.5),
  domain = rep(c("Arithmetic", "Algebra"), each = 3),
  stringsAsFactors = FALSE
)
```

---

# Part 2: The page flow

`custom_page_flow` is a plain R list. Each element describes one page. **The order of the list is the order participants see.** To reorder pages, reorder the list.

## 2.1 Four page types

| `type` | What participants see |
|----------------|-------------------------------------------------------|
| `"custom"` | Free HTML — intro, consent, transition text, or debriefing. |
| `"demographics"` | A form built automatically from `demographic_configs`. |
| `"items"` | Questionnaire items drawn from the item bank by row index. |
| `"results"` | The final screen — built-in score table and/or a custom results function. |

Every page entry needs `id`, `type`, `title`, and `title_en`.

## 2.2 Defining pages

Here we define four page objects. Nothing renders yet; we combine them in Part 4.

```{r page_examples}
# Welcome screen
intro_page <- list(
  id         = "intro",
  type       = "custom",
  title      = "Welcome",
  title_en   = "Welcome",
  content    = "<h2>Welcome</h2><p>This study takes about 5 minutes.</p>",
  content_en = "<h2>Welcome</h2><p>This study takes about 5 minutes.</p>"
)

# Background questions — inrep builds the form from demographic_configs
demos_page <- list(
  id           = "demos",
  type         = "demographics",
  title        = "About you",
  title_en     = "About you",
  demographics = c("age", "gender"),   # IDs from demographic_configs
  required     = TRUE
)

# Items — reference rows by position
items_page <- list(
  id              = "block_1",
  type            = "items",
  title           = "Questions",
  title_en        = "Questions",
  item_indices    = 1:6,              # rows 1-6 of the item bank
  scale_type      = "likert",
  instructions    = "Rate how well each statement describes you.",
  instructions_en = "Rate how well each statement describes you."
)

# Results — built-in score table with declarative scoring
results_page <- list(
  id                = "results",
  type              = "results",
  title             = "Your results",
  title_en          = "Your results",
  show_scale_scores = TRUE,
  results_text_en   = "Thank you for participating.",
  report_metrics    = list(
    list(name = "Conscientiousness", label = "Conscientiousness",
         items = paste0("C", 1:6), formula = "mean")
  )
)
```

`report_metrics` is the simplest scoring path: list which item IDs belong to each scale and inrep computes the means and shows a built-in table. For custom charts, reverse-coded scoring, or richer HTML feedback, use a `results_processor` function — covered in Vignette 2.

> **Robust indices.** `item_indices = 1:6` points to rows, not item IDs. If items are added or reordered, those positions shift. Compute from IDs or domain labels instead so they stay correct after any reordering:
>
> ``` r
> c_idx <- which(my_items$id %in% paste0("C", 1:6))
> e_idx <- which(bfi_items$domain == "Extraversion")
> ```

---

# Part 3: Demographics

`demographic_configs` is a named list — one entry per background question. Entry names are what you reference in page definitions. Field types: `"number"`, `"select"`, `"checkbox"`, `"slider"`, `"text"`.

When we run this chunk, we create a named list of demographic fields. It is used later in `create_study_config()`.

```{r demo_configs}
demographic_configs <- list(

  age = list(
    question    = "How old are you?",
    question_en = "How old are you?",
    type        = "number",
    required    = TRUE
  ),

  gender = list(
    question    = "What is your gender?",
    question_en = "What is your gender?",
    type        = "select",
    options    = c("Female" = "f", "Male" = "m",
                   "Diverse" = "d", "Prefer not to say" = "na"),
    options_en = c("Female" = "f", "Male" = "m",
                   "Diverse" = "d", "Prefer not to say" = "na"),
    required   = TRUE
  ),

  stress = list(
    question     = "How stressed do you feel right now?",
    question_en  = "How stressed do you feel right now?",
    type         = "slider",
    min = 0, max = 100, step = 1, default = 50,
    label_min    = "Not at all",  label_max    = "Extremely",
    label_min_en = "Not at all",  label_max_en = "Extremely",
    required     = TRUE
  ),

  show_results = list(
    question    = "Would you like to see your personal results at the end?",
    question_en = "Would you like to see your personal results at the end?",
    type        = "select",
    options    = c("Yes" = "yes", "No" = "no"),
    options_en = c("Yes" = "yes", "No"  = "no"),
    required   = TRUE
  )
)
```

> **Slider values** are stored in the demographics output, not the responses vector. Reading a slider field from `responses[30]` returns `NA` silently. Vignette 2 covers reading demographic values.

`create_study_config()` takes two related arguments: `demographics` (the field IDs to activate) and `demographic_configs` (the full definitions). Pass both:

This snippet shows exactly where the two demographics arguments go in `create_study_config()`.

```{r demos_wiring}
study_config <- create_study_config(
  # ... other arguments ...
  demographics        = names(demographic_configs),
  demographic_configs = demographic_configs
)
```

**Bilingual studies:** set `bilingual = TRUE` and pair every participant-facing string with an `_en` variant. inrep adds a language toggle automatically. Do not implement language switching via custom JavaScript — Shiny re-renders reset JS state.

---

# Part 3b: Themes

Pass one theme name to `create_study_config(theme = ...)`. That is all that is needed.

**Core themes:** `"Light"`, `"Midnight"`, `"Sunset"`, `"Forest"`, `"Ocean"`, `"Berry"`, `"Professional"`, `"hildesheim"`

**Additional themes** — inspired by popular AI tool interfaces and sourced from their public design systems: `"claude"`, `"opencode"`, `"xai"`, `"mistral"`, `"notion"`, `"linear"`, `"vercel"`, `"stripe"`, `"raycast"`, `"warp"`, `"paper"`, `"sepia"`, `"dark-mode"`, `"high-contrast"`, `"colorblind-safe"`, `"dyslexia-friendly"`, `"large-text"`, and more.

```{r themes}
# Pass any theme name as a string — the complete study below uses "hildesheim"
create_study_config(name = "Demo", theme = "hildesheim", adaptive = FALSE)
create_study_config(name = "Demo", theme = "claude",     adaptive = FALSE)
create_study_config(name = "Demo", theme = "Light",      adaptive = FALSE)
```

---

# Part 4: A complete study

Everything above, assembled into one script. This example follows the pattern of the **Hildesheim study** (`case_studies/hildesheim_study/HilFo.R`) — the most fully-featured study in the package: bilingual DE/EN, the `"hildesheim"` theme, `session_save = TRUE` for resumable sessions, and a results page that can show or hide scores per participant.

`launch_study()` starts a local server and opens the study in your default browser. The R console shows `Listening on http://127.0.0.1:XXXX` and stays busy until you stop it (Stop button in RStudio, or Escape in the console).

With `session_save = TRUE` and `report_formats = c("csv", "json")`, responses are written to `study_data/` in your working directory.

This getting-started example keeps demographic fields optional so the flow never blocks at page 2 during a first test run. In production studies, set required fields back to `TRUE`.

**Debug shortcuts** (only when `debug_mode = TRUE`, red indicator bottom-right): **Ctrl+A** fills the current page; **Ctrl+Q** fills all pages and jumps straight to the results screen.

When we run this chunk, the study opens in your browser and the R console shows `Listening on http://127.0.0.1:XXXX`. During testing, use **Ctrl+Q** to jump to results. Set `debug_mode = FALSE` before sharing with participants.

```{r complete_study}
library(inrep)

# ============================================================
# 1. Item bank
# ============================================================
data(bfi_items)

# Add stable IDs, domain labels, and reverse-coding flags
bfi_items$id <- paste0("BFI_", sprintf("%02d", seq_len(nrow(bfi_items))))
bfi_items$domain <- rep(
  c("Extraversion", "Agreeableness", "Conscientiousness", "Neuroticism", "Openness"),
  each = 6
)
bfi_items$reverse_coded <- rep(c(FALSE, TRUE, FALSE, TRUE, FALSE, FALSE), times = 5)

# ============================================================
# 2. Demographics
# ============================================================
demographic_configs <- list(

  consent = list(
    question    = "I agree to participate voluntarily and anonymously.",
    question_en = "I agree to participate voluntarily and anonymously.",
    type        = "checkbox",
    required    = FALSE
  ),

  age = list(
    question    = "How old are you?",
    question_en = "How old are you?",
    type        = "number",
    required    = FALSE
  ),

  gender = list(
    question    = "What is your gender?",
    question_en = "What is your gender?",
    type        = "select",
    options    = c("Female" = "f", "Male" = "m",
                   "Diverse" = "d", "Prefer not to say" = "na"),
    options_en = c("Female" = "f", "Male" = "m",
                   "Diverse" = "d", "Prefer not to say" = "na"),
    required   = FALSE
  ),

  show_results = list(
    question    = "Would you like to see your personality profile at the end?",
    question_en = "Would you like to see your personality profile at the end?",
    type        = "select",
    options    = c("Yes" = "yes", "No" = "no"),
    options_en = c("Yes" = "yes", "No"  = "no"),
    required   = FALSE
  )
)

# ============================================================
# 3. Page flow
# Compute indices from domain labels — stays correct if bfi_items is reordered
# ============================================================
ext_idx <- which(bfi_items$domain == "Extraversion")
neu_idx <- which(bfi_items$domain == "Neuroticism")

custom_page_flow <- list(

  list(id = "intro",   type = "custom",
       title = "Welcome", title_en = "Welcome",
       content = paste0(
         "<h2>Welcome to our study</h2>",
         "<p>We are examining personality and wellbeing. ",
         "Participation is voluntary and anonymous. ",
         "The questionnaire takes about 5 minutes.</p>"
       ),
       content_en = paste0(
         "<h2>Welcome to our study</h2>",
         "<p>We are examining personality and wellbeing. ",
         "Participation is voluntary and anonymous. ",
         "The questionnaire takes about 5 minutes.</p>"
       )),

  list(id = "demos",   type = "demographics",
       title = "About you", title_en = "About you",
       demographics = names(demographic_configs),
      required = FALSE),

  list(id = "block_1", type = "items",
       title = "Part 1 of 2", title_en = "Part 1 of 2",
       item_indices    = ext_idx,
       scale_type      = "likert",
       instructions    = "How well does each statement describe you?",
       instructions_en = "How well does each statement describe you?"),

  list(id = "block_2", type = "items",
       title = "Part 2 of 2", title_en = "Part 2 of 2",
       item_indices    = neu_idx,
       scale_type      = "likert",
       instructions    = "How well does each statement describe you?",
       instructions_en = "How well does each statement describe you?"),

  list(id = "results", type = "results",
       title    = "Your profile",
       title_en = "Your profile",
       show_scale_scores = TRUE,
       show_radar_chart  = FALSE,
       results_text_en   = "Thank you for your participation.",
       report_metrics    = list(
         list(name = "Extraversion", label = "Extraversion",
              items = bfi_items$id[bfi_items$domain == "Extraversion"],
              formula = "mean"),
         list(name = "Neuroticism",  label = "Neuroticism",
              items = bfi_items$id[bfi_items$domain == "Neuroticism"],
              formula = "mean")
       ))
)

# ============================================================
# 4. Study config  (Hildesheim-style: bilingual, branded theme)
# ============================================================
study_config <- create_study_config(
  name                = "Personality Study – Hildesheim",
  custom_page_flow    = custom_page_flow,
  demographics        = names(demographic_configs),
  demographic_configs = demographic_configs,
  model               = "GRM",
  adaptive            = FALSE,
  min_items           = length(c(ext_idx, neu_idx)),
  max_items           = length(c(ext_idx, neu_idx)),
  # Hildesheim-style settings:
  theme               = "hildesheim",   # University of Hildesheim brand colours
  language            = "de",           # primary language German
  bilingual           = TRUE,           # adds DE/EN toggle to every page
  response_ui_type    = "radio",
  progress_style      = "bar",
  session_save        = TRUE,           # sessions can be resumed if interrupted
  report_formats      = c("csv", "json"),
  show_scale_scores   = TRUE            # set FALSE to show offboarding only
)

# ============================================================
# 5. Launch
# debug_mode = TRUE: skip through the study instantly during testing.
# Set to FALSE before sharing the study URL.
# ============================================================
launch_study(config = study_config, item_bank = bfi_items, debug_mode = TRUE)
```
