---
title: "Scoring and Reporting in inrep"
author: "Clievins Selva"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: cerulean
vignette: >
  %\VignetteIndexEntry{Scoring and Reporting in inrep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(inrep)
data(bfi_items)

# bfi_items has no id, domain or keying column, so we add them here. The items
# are not ordered by domain; the assignment below follows the item content.
bfi_items$id <- paste0("BFI_", sprintf("%02d", seq_len(nrow(bfi_items))))
bfi_domains <- list(
  Extraversion      = c(1, 2, 3, 14, 18, 22, 25),
  Agreeableness     = c(4, 5, 11, 12, 13, 19, 26),
  Conscientiousness = c(6, 17, 24, 27, 30),
  Neuroticism       = c(7, 8, 16, 21, 23, 28),
  Openness          = c(9, 10, 15, 20, 29)
)
bfi_items$domain <- NA_character_
for (d in names(bfi_domains)) bfi_items$domain[bfi_domains[[d]]] <- d
# reserved, rude, emotionally stable, starts quarrels, shy, relaxed, calm
bfi_items$reverse_coded <- seq_len(nrow(bfi_items)) %in% c(2, 5, 8, 12, 14, 21, 28)

# One simulated participant for all previews below
set.seed(42)
responses    <- sample(1:5, nrow(bfi_items), replace = TRUE)
demographics <- list(
  age                   = 24,
  gender                = "f",
  stress_level          = 65,
  show_personal_results = "yes"
)

# Mean of a scale after reverse coding; NA when fewer than two items were answered
score_scale <- function(ids, responses, item_bank, max_response = 5) {
  idx <- match(ids, item_bank$id)
  raw <- responses[idx]
  rev <- item_bank$reverse_coded[idx]
  raw[rev] <- (max_response + 1) - raw[rev]
  if (sum(!is.na(raw)) < 2) NA_real_ else mean(raw, na.rm = TRUE)
}

# Used in Part 7
safe_sd <- function(x) { v <- x[!is.na(x)]; if (length(v) < 2) NA_real_ else sd(v) }

# Simulated reference sample for Part 6 and the norm curve in Part 3
set.seed(100)
norm_pop <- data.frame(
  Extraversion      = pmin(5, pmax(1, rnorm(300, 3.2, 0.7))),
  Agreeableness     = pmin(5, pmax(1, rnorm(300, 3.7, 0.6))),
  Conscientiousness = pmin(5, pmax(1, rnorm(300, 3.5, 0.6))),
  Neuroticism       = pmin(5, pmax(1, rnorm(300, 2.9, 0.8))),
  Openness          = pmin(5, pmax(1, rnorm(300, 3.4, 0.6)))
)
pct_rank <- function(score, reference) {
  if (is.na(score) || length(reference) < 5) return(NA_real_)
  round(mean(reference < score, na.rm = TRUE) * 100)
}

# Writes a preview into the page. Line breaks are removed first, since pandoc
# would otherwise read indented HTML lines as code blocks.
show_html <- function(x) cat(gsub("\\s*\n\\s*", " ", as.character(x)))

# Scores of the simulated participant
dims   <- c("Extraversion", "Agreeableness", "Conscientiousness", "Neuroticism", "Openness")
scores <- setNames(lapply(dims, function(d)
  score_scale(bfi_items$id[bfi_items$domain == d], responses, bfi_items)), dims)
```

# Why a results page

Many participants, in particular those who take part for course credit, have little reason to read every item carefully. Attention checks and response-time thresholds can remove careless responses afterwards, but they reduce the sample and do not change why people responded carelessly in the first place.

A results page approaches the problem from the other side. If participants know that they will receive something in return, a profile of their own scores or at least a confirmation that their answers arrived, the survey becomes less of a one-sided exchange. We use this in the HilFo study (`case_studies/hildesheim_study/HilFo.R`), where first-semester students see their Big Five, stress and study skill scores at the end. This vignette describes how inrep builds such a page, starting from a plain table and moving on to charts.

------------------------------------------------------------------------

# Two optional functions

The report is controlled by two functions, both optional and independent of each other.

| Function | Called when | Returns | Read by |
|---|---|---|---|
| `results_processor` | the participant reaches a results page | `shiny::HTML()` | the participant |
| `csv_processor` | the participant downloads a CSV from the report | the modified `csv_data` | the participant |

Without a `results_processor`, participants see inrep's default results page. The `csv_processor` only shapes the file participants can download; it does not change the data inrep stores for the researcher (Part 5 explains how to add scale scores to those).

------------------------------------------------------------------------

# Part 1: The results_processor

## 1.1 What the function receives

```{r p1_sig, eval = FALSE}
my_report <- function(responses, item_bank, demographics = NULL,
                      session = NULL, rv = NULL) {
  # compute scores, build the HTML
  shiny::HTML("<p>...</p>")
}
```

inrep looks at the argument names and passes only those the function asks for. `responses` and `item_bank` are always passed.

- `responses` is numeric and follows the row order of `item_bank`. In adaptive studies, or when a participant skipped the last pages, it can be shorter than `nrow(item_bank)`, so we pad it before scoring.
- `item_bank` is the data frame given to `launch_study()`. We look up items by `id` rather than by position.
- `demographics` is a named list with everything from the demographics pages, including sliders, which are not part of `responses`.
- `session` is the Shiny session on the final results page.
- `rv` holds the state of the running study. `rv$language` is the language the participant currently sees, which a bilingual report needs (Part 4).

`input`, `page` and `config` can be requested in the same way. When testing a processor outside a running study, all optional arguments are simply `NULL`.

The function has to return `shiny::HTML()`. A plain character string is shown as text, with the HTML tags visible.

## 1.2 A minimal working example

```{r p1_minimal}
minimal_report <- function(responses, item_bank,
                           demographics = NULL, session = NULL) {
  n <- nrow(item_bank)
  if (length(responses) < n)
    responses <- c(responses, rep(NA, n - length(responses)))
  responses <- as.numeric(responses)

  overall <- round(mean(responses, na.rm = TRUE), 2)
  n_done  <- sum(!is.na(responses))

  shiny::HTML(paste0(
    "<div style='padding:40px;text-align:center;font-family:sans-serif;'>",
    "<h2>Thank you for participating!</h2>",
    "<p>You completed <strong>", n_done, "</strong> items.</p>",
    "<p>Overall mean response: <strong>", overall, "</strong> (scale 1-5)</p>",
    "<p style='color:#888;font-size:0.9em;'>Your responses have been saved.</p>",
    "</div>"
  ))
}
```

This is what the participant sees on the last page:

```{r p1_preview, results = "asis", echo = FALSE}
cat('<div style="border:2px solid #4A90D9;border-radius:6px;margin:20px 0;overflow:hidden;">')
cat('<div style="background:#4A90D9;color:white;padding:6px 16px;font-size:12px;font-weight:bold;letter-spacing:.5px;">PARTICIPANT VIEW</div>')
show_html(minimal_report(responses, bfi_items, demographics, NULL))
cat('</div>')
```

## 1.3 Passing it to the study

```{r p1_wire, eval = FALSE}
config <- inrep::create_study_config(
  name              = "BFI pilot",
  model             = "GRM",
  max_items         = 10,
  results_processor = minimal_report   # the function itself, not its name
)
inrep::launch_study(config = config, item_bank = bfi_items, debug_mode = TRUE)
```

## 1.4 Study level and page level

The processor given to `create_study_config()` is used on every results page. A results page in `custom_page_flow` can set its own `results_processor`, which then applies to that page only. On the page it may also be the name of a function (`"minimal_report"`), which is useful when the page flow is defined before the function, as in HilFo. inrep looks the name up in the global environment and falls back to the study-level processor if no such function exists.

------------------------------------------------------------------------

# Part 2: Scoring

## 2.1 Scoring by item ID

Since `responses` follows the row order of the item bank, it is tempting to score by position (`responses[1:6]`). This works until items are added or reordered, and in an adaptive study it does not work at all. We therefore look items up by their `id`, which stays the same however the rows are arranged.

`score_scale()` in the setup chunk of this vignette does this: it finds the items by ID, reverses the items flagged in `reverse_coded`, and returns the mean. A scale with fewer than two answered items is left missing. For the simulated participant:

```{r p2_scores}
round(unlist(scores), 2)
```

## 2.2 Reverse coding

A reverse-coded item is recoded as `(max + 1) - response`, so on a 1-5 scale a 2 becomes a 4. This has to happen before the items are averaged. `score_scale()` takes the information from the `reverse_coded` column, so the keying is defined once in the item bank and not again in every function that scores.

## 2.3 Sliders are not in `responses`

Visual analogue sliders (0-100) are defined as demographic questions in inrep, so their values arrive in `demographics` and not in `responses`. To report them next to Likert items, we put them on the same 1-5 metric:

```{r p2_slider}
slider_raw <- as.numeric(demographics$stress_level)   # 0-100
stat_score <- round((slider_raw / 100) * 4 + 1, 2)   # rescale 0-100 to 1-5
stat_score
```

With `start_empty = TRUE`, a slider has no handle until the participant clicks on it, and inrep additionally sends `demo_<name>_touched`. HilFo uses this flag so that a slider nobody moved is stored as missing rather than as its starting value (see `slider_value()` in `HilFo.R`).

## 2.4 A frame for every processor

If the processor fails, inrep catches the error, keeps the data and shows a short error message in English that includes the R error text. That is fine while testing, but participants should rather see a message in their own language. We therefore wrap the processor in its own `tryCatch()` and pad the responses before scoring, since a shorter vector would otherwise shift the item positions without any warning.

```{r p2_skeleton, eval = FALSE}
my_report <- function(responses, item_bank, demographics = NULL, rv = NULL) {
  is_english <- is.null(rv) || identical(shiny::isolate(rv$language), "en")
  tryCatch({
    n <- nrow(item_bank)
    if (length(responses) < n)
      responses <- c(responses, rep(NA, n - length(responses)))
    responses <- as.numeric(responses)

    # score and build the HTML here

    shiny::HTML("<div>...</div>")
  }, error = function(e) {
    message("Report failed: ", conditionMessage(e))
    shiny::HTML(if (is_english)
      "<p>Your answers were saved, but the report could not be shown.</p>"
    else
      "<p>Ihre Antworten wurden gespeichert, die Auswertung konnte aber nicht angezeigt werden.</p>")
  })
}
```

`rv` is `NULL` when the function is called outside a running study, so the report then defaults to English. `shiny::isolate()` reads the language without making the report depend on it reactively.

------------------------------------------------------------------------

# Part 3: Visualizing scores

We show five ways to present the scores, ordered from the simplest to the most flexible. Our recommendation is to use the simplest one that answers the participant's question. A table is easy to check against the data, works with screen readers and cannot be misread because of an axis; a chart is worth adding when it shows something the table does not, such as the shape of a profile.

The helper functions below are not part of inrep. They are written out here so that they can be copied into a study script, or into a separate file that is `source()`d before `launch_study()`.

| Presentation | Needs | Suited for |
|---|---|---|
| HTML table | nothing | means and standard deviations |
| HTML/CSS bars | nothing | scores relative to the scale maximum |
| SVG radar | nothing | the shape of a profile across three or more scales |
| SVG norm curve | nothing | the position of a score in a reference sample |
| ggplot as PNG | ggplot2, base64enc | anything ggplot2 can draw |

All examples use the `scores` from Part 2.

## 3.1 HTML table

We would include a table in every report, even next to charts, since it is the one format that survives screenshots, screen readers and copying.

```{r p3_table_fn}
score_table_html <- function(scores, lang = "en") {
  rows <- vapply(names(scores), function(k) {
    v <- if (is.na(scores[[k]])) "&mdash;" else sprintf("%.2f", scores[[k]])
    sprintf(
      "<tr>
         <td style='padding:10px;border-bottom:1px solid #eee;'>%s</td>
         <td style='padding:10px;text-align:center;font-weight:bold;'>%s</td>
       </tr>", k, v)
  }, character(1))

  paste0(
    "<table style='width:100%;max-width:480px;margin:8px auto;border-collapse:collapse;",
    "font-family:system-ui,sans-serif;'>",
    "<thead><tr style='background:#f8fafc;border-bottom:2px solid #e5e7eb;'>",
    "<th style='padding:10px;text-align:left;'>",
    if (lang == "en") "Scale" else "Skala",
    "</th>",
    "<th style='padding:10px;text-align:center;'>",
    if (lang == "en") "Mean" else "Wert",
    "</th></tr></thead>",
    "<tbody>", paste(rows, collapse = ""), "</tbody>",
    "</table>"
  )
}
```

```{r p3_table_preview, results = "asis", echo = FALSE}
cat('<div style="border:2px solid #4A90D9;border-radius:6px;margin:20px 0;overflow:hidden;">')
cat('<div style="background:#4A90D9;color:white;padding:6px 16px;font-size:12px;font-weight:bold;">PARTICIPANT VIEW: score table</div>')
show_html(score_table_html(scores))
cat('</div>')
```

## 3.2 HTML/CSS progress bars

A bar is simply a `<div>` inside another `<div>`, with the width of the inner one set to the score as a percentage of the scale maximum. No package is needed, and the bars scale with the screen.

```{r p3_bars_fn}
score_row_with_bar <- function(label, score, sd_val = NA_real_,
                               max_response = 5, color = "#4A90D9") {
  pct      <- if (is.na(score)) 0 else round(100 * score / max_response)
  m_str    <- if (is.na(score))  "&mdash;" else sprintf("%.2f", score)
  sd_str   <- if (is.na(sd_val)) "&mdash;" else sprintf("%.2f", sd_val)

  sprintf(
    "<tr style='border-bottom:1px solid #f3f4f6;'>
       <td style='padding:10px 14px;font-weight:500;'>%s</td>
       <td style='padding:10px 14px;text-align:center;color:%s;font-weight:600;'>%s</td>
       <td style='padding:10px 14px;text-align:center;color:#6b7280;'>%s</td>
       <td style='padding:10px 14px;'>
         <div style='background:#e5e7eb;border-radius:4px;height:8px;overflow:hidden;'>
           <div style='background:%s;height:100%%;width:%d%%;border-radius:4px;'></div>
         </div>
       </td>
     </tr>",
    label, color, m_str, sd_str, color, pct)
}

scores_bar_table <- function(scores, sds = NULL, lang = "en") {
  if (is.null(sds)) sds <- setNames(rep(NA_real_, length(scores)), names(scores))
  rows <- vapply(names(scores),
    function(k) score_row_with_bar(k, scores[[k]], sds[[k]]),
    character(1))

  paste0(
    "<div style='background:#fff;border:1px solid #e5e7eb;border-radius:8px;",
    "overflow:hidden;max-width:640px;margin:8px auto;'>",
    "<table style='width:100%;border-collapse:collapse;font-family:system-ui,sans-serif;'>",
    "<thead><tr style='background:#f9fafb;border-bottom:2px solid #e5e7eb;'>",
    "<th style='padding:12px 14px;text-align:left;'>",
    if (lang == "en") "Scale" else "Skala", "</th>",
    "<th style='padding:12px 14px;text-align:center;'>M</th>",
    "<th style='padding:12px 14px;text-align:center;'>SD</th>",
    "<th style='padding:12px 14px;width:35%;'></th>",
    "</tr></thead><tbody>", paste(rows, collapse = ""), "</tbody></table>",
    "<div style='padding:8px 14px;background:#f9fafb;border-top:1px solid #e5e7eb;",
    "font-size:0.75rem;color:#9ca3af;'><em>",
    if (lang == "en") "M = Mean. Scale: 1-5." else "M = Mittelwert. Skala: 1-5.",
    "</em></div></div>"
  )
}
```

```{r p3_bars_preview, results = "asis", echo = FALSE}
sds_example <- setNames(
  sapply(dims, function(d)
    sd(responses[match(bfi_items$id[bfi_items$domain == d], bfi_items$id)], na.rm = TRUE)),
  dims)
cat('<div style="border:2px solid #27AE60;border-radius:6px;margin:20px 0;overflow:hidden;">')
cat('<div style="background:#27AE60;color:white;padding:6px 16px;font-size:12px;font-weight:bold;">PARTICIPANT VIEW: M/SD bar table</div>')
show_html(scores_bar_table(scores, sds_example))
cat('</div>')
```

## 3.3 Inline SVG radar

With three or more scales, a radar chart shows the shape of a profile, which bars show less directly. The chart below is built as an SVG string, again without any package.

```{r p3_radar_fn}
svg_radar <- function(scores, max_response = 5, color = "#4A90D9",
                      track = "#e5e7eb", text_color = "#374151", size = 280) {
  vals   <- unlist(scores)
  labels <- names(scores)[!is.na(vals)]
  vals   <- vals[!is.na(vals)]
  if (length(vals) < 3) return("")   # need at least 3 axes

  n  <- length(labels)
  vb <- 240; cx <- 120; cy <- 120; r <- 85
  angles <- seq(0, 2 * pi, length.out = n + 1)[1:n] - pi / 2

  bg <- paste(sprintf(
    "<circle cx='%g' cy='%g' r='%g' fill='none' stroke='%s' stroke-width='1'/>",
    cx, cy, r * c(1, 0.66, 0.33), track), collapse = "")

  axes <- vapply(seq_along(angles), function(i) {
    x2 <- cx + r * cos(angles[i]); y2 <- cy + r * sin(angles[i])
    lx <- cx + (r + 22) * cos(angles[i]); ly <- cy + (r + 22) * sin(angles[i])
    sprintf(
      "<line x1='%g' y1='%g' x2='%g' y2='%g'
             stroke='%s' stroke-width='1' stroke-dasharray='3,3'/>
       <text x='%g' y='%g' text-anchor='middle'
             font-family='system-ui' font-size='11'
             fill='%s' dominant-baseline='middle'>%s</text>",
      cx, cy, x2, y2, track, lx, ly, text_color, labels[i])
  }, character(1))

  vn  <- vals / max_response
  pts <- paste(sprintf("%g,%g",
    cx + r * vn * cos(angles), cy + r * vn * sin(angles)), collapse = " ")
  dots <- vapply(seq_along(angles), function(i)
    sprintf("<circle cx='%g' cy='%g' r='4' fill='%s' stroke='white' stroke-width='1.5'/>",
            cx + r * vn[i] * cos(angles[i]), cy + r * vn[i] * sin(angles[i]), color),
    character(1))

  rgb_str <- paste(col2rgb(color), collapse = ",")
  paste0(
    # the margin around the 240 x 240 plot leaves room for long axis labels
    "<svg viewBox='-50 -10 ", vb + 100, " ", vb + 20, "' xmlns='http://www.w3.org/2000/svg' ",
    "style='width:100%;max-width:", size, "px;display:block;margin:16px auto;'>",
    bg, paste(axes, collapse = ""),
    sprintf("<polygon points='%s' fill='rgba(%s,0.22)' stroke='%s'
                     stroke-width='2' stroke-linejoin='round'/>", pts, rgb_str, color),
    paste(dots, collapse = ""), "</svg>")
}
```

With fewer than three scored scales the function returns an empty string, since a radar chart with one or two axes says nothing a table would not say better.

```{r p3_radar_preview, results = "asis", echo = FALSE}
cat('<div style="border:2px solid #E74C3C;border-radius:6px;margin:20px 0;overflow:hidden;">')
cat('<div style="background:#E74C3C;color:white;padding:6px 16px;font-size:12px;font-weight:bold;">PARTICIPANT VIEW: SVG radar chart</div>')
show_html(svg_radar(scores))
cat('</div>')
```

## 3.4 Inline SVG norm curve

A normal curve with the participant's score marked and the area to its left shaded shows where the score lies in a reference sample.

```{r p3_normcurve_fn}
svg_norm_curve <- function(score, norm_mean, norm_sd,
                           label       = "Reference population",
                           color       = "#4A90D9",
                           shade_color = "rgba(74,144,217,0.20)",
                           max_response = 5,
                           width = 480, height = 130) {
  if (is.na(score) || is.na(norm_mean) || is.na(norm_sd) || norm_sd <= 0) return("")

  # +/- 3.5 SD around the mean, but not beyond the ends of the scale
  x_lo <- max(1, norm_mean - 3.5 * norm_sd)
  x_hi <- min(max_response, norm_mean + 3.5 * norm_sd)
  score <- max(x_lo, min(x_hi, score))   # clip score to drawn range

  pad_l <- 28; pad_r <- 28; pad_t <- 22; pad_b <- 36
  plot_w <- width  - pad_l - pad_r
  plot_h <- height - pad_t - pad_b

  to_x <- function(v) pad_l + plot_w * (v - x_lo) / (x_hi - x_lo)
  max_d <- dnorm(norm_mean, norm_mean, norm_sd)
  to_y  <- function(d) height - pad_b - plot_h * (d / max_d)

  # 80-point curve
  xs <- seq(x_lo, x_hi, length.out = 80)
  ds <- dnorm(xs, norm_mean, norm_sd)
  pts <- paste(sprintf("%g,%g", to_x(xs), to_y(ds)), collapse = " ")

  # Shaded area up to the participant's score
  xs_s <- c(xs[xs < score], score)
  ds_s <- c(dnorm(xs_s[-length(xs_s)], norm_mean, norm_sd), dnorm(score, norm_mean, norm_sd))
  base_y <- height - pad_b
  shade_pts <- paste(c(
    sprintf("%g,%g", to_x(xs_s[1]),  base_y),
    sprintf("%g,%g", to_x(xs_s),     to_y(ds_s)),
    sprintf("%g,%g", to_x(xs_s[length(xs_s)]), base_y)
  ), collapse = " ")

  # Axis ticks
  tick_vals <- seq(ceiling(x_lo), floor(x_hi))
  ticks <- vapply(tick_vals, function(k) {
    sprintf(
      "<line x1='%g' y1='%g' x2='%g' y2='%g' stroke='#9ca3af' stroke-width='1'/>
       <text x='%g' y='%g' text-anchor='middle' font-family='system-ui'
             font-size='10' fill='#6b7280'>%g</text>",
      to_x(k), base_y, to_x(k), base_y + 4,
      to_x(k), base_y + 16, k)
  }, character(1))

  # Participant marker: vertical line + dot + score label
  px  <- to_x(score)
  pct <- round(pnorm(score, norm_mean, norm_sd) * 100)
  marker <- sprintf(
    "<line x1='%g' y1='%g' x2='%g' y2='%g' stroke='%s' stroke-width='2' stroke-dasharray='3,3'/>
     <circle cx='%g' cy='%g' r='5' fill='%s' stroke='white' stroke-width='1.5'/>
     <text x='%g' y='%g' text-anchor='middle' font-family='system-ui'
           font-size='11' font-weight='600' fill='%s'>%.2f (%d%%)</text>",
    px, to_y(max_d * 0.08), px, base_y, color,
    px, base_y, color,
    px, pad_t + 2, color, score, pct)

  # Label below the chart
  lbl <- sprintf(
    "<text x='%g' y='%g' text-anchor='middle' font-family='system-ui'
           font-size='10' fill='#9ca3af'>%s (M = %.2f, SD = %.2f)</text>",
    width / 2, height - 2, label, norm_mean, norm_sd)

  paste0(
    "<svg viewBox='0 0 ", width, " ", height, "' xmlns='http://www.w3.org/2000/svg'",
    " style='width:100%;max-width:", width, "px;display:block;margin:12px auto;'>",
    # Baseline
    sprintf("<line x1='%g' y1='%g' x2='%g' y2='%g' stroke='#e5e7eb' stroke-width='1'/>",
            to_x(x_lo), base_y, to_x(x_hi), base_y),
    # Shaded area
    "<polygon points='", shade_pts, "' fill='", shade_color, "'/>",
    # Curve
    "<polyline points='", pts, "' fill='none' stroke='", color,
    "' stroke-width='2' stroke-linecap='round'/>",
    # Ticks
    paste(ticks, collapse = ""),
    # Marker
    marker, lbl,
    "</svg>"
  )
}
```

```{r p3_normcurve_preview, results = "asis", echo = FALSE}
cat('<div style="border:2px solid #8E44AD;border-radius:6px;margin:20px 0;overflow:hidden;">')
cat('<div style="background:#8E44AD;color:white;padding:6px 16px;font-size:12px;font-weight:bold;">PARTICIPANT VIEW: SVG norm curve (Extraversion)</div>')
norm_E <- norm_pop$Extraversion
show_html(svg_norm_curve(scores$Extraversion, mean(norm_E), sd(norm_E),
                   label = "Reference population (N = 300)"))
cat('</div>')
```

The shaded area and the percentage at the marker come from a normal distribution with the mean and standard deviation of the reference sample. They are an approximation of the percentile rank: when the reference scores are skewed or bounded by the scale, the empirical rank from `pct_rank()` (Part 6) can differ noticeably, and that is the number we would report in a table.

## 3.5 Embedding a ggplot

For anything the SVG helpers do not cover, such as boxplots, densities or scatter plots, a ggplot can be saved to a temporary PNG and embedded as base64. This is what HilFo does for its radar and bar charts. Each PNG adds tens of kilobytes to the page, while the SVG functions above add a few, so with several charts on one page SVG loads noticeably faster on a phone.

```{r p3_ggplot, eval = FALSE}
gg_to_html <- function(p, width = 8, height = 4, dpi = 150) {
  if (!requireNamespace("ggplot2",   quietly = TRUE)) return("")
  if (!requireNamespace("base64enc", quietly = TRUE)) return("")
  f <- tempfile(fileext = ".png")
  on.exit(unlink(f), add = TRUE)
  ggplot2::ggsave(f, p, width = width, height = height, dpi = dpi, bg = "white")
  b64 <- base64enc::base64encode(f)
  sprintf(
    "<img src='data:image/png;base64,%s' alt='chart'
     style='width:100%%;max-width:760px;display:block;margin:16px auto;'>", b64)
}

# inside the results_processor
sdf <- data.frame(dim = names(scores), score = unlist(scores))
p   <- ggplot2::ggplot(sdf, ggplot2::aes(x = dim, y = score)) +
  ggplot2::geom_col(fill = "#4A90D9", width = 0.6) +
  ggplot2::geom_hline(yintercept = 3, linetype = "dashed", color = "#aaa") +
  ggplot2::scale_y_continuous(limits = c(0, 5.5), breaks = 1:5) +
  ggplot2::labs(x = NULL, y = "Score (1-5)") +
  ggplot2::theme_minimal(base_size = 13)

img_html <- gg_to_html(p)
```

------------------------------------------------------------------------

# Part 4: Branching the report

## 4.1 Privacy opt-out

Some participants are happy to contribute their data but do not want to see their own scores. HilFo asks about this on the page before the report (`show_personal_results`). The answer is checked before anything is scored or plotted, since there is no reason to build charts nobody will see.

```{r p4_privacy}
privacy_aware_report <- function(responses, item_bank,
                                 demographics = NULL, rv = NULL) {
  is_english <- is.null(rv) || identical(shiny::isolate(rv$language), "en")

  opted_out <- isTRUE(
    tolower(as.character(demographics$show_personal_results)) %in%
      c("no", "n", "false", "0")
  )
  if (opted_out) {
    return(shiny::HTML(
      if (is_english)
        "<div style='padding:40px;text-align:center;font-family:sans-serif;'>
           <h2>Assessment complete</h2>
           <p>Thank you. Your data has been saved.</p></div>"
      else
        "<div style='padding:40px;text-align:center;font-family:sans-serif;'>
           <h2>Vielen Dank!</h2>
           <p>Ihre Daten wurden gespeichert.</p></div>"
    ))
  }

  # the participant wants the report
  shiny::HTML("<p>... personalized report here ...</p>")
}
```

## 4.2 Two languages

In a bilingual study the report follows the language the participant chose, which inrep keeps in `rv$language`. We collect all texts of the report in one list `L` at the top of the function and use `L$title`, `L$note` and so on further down. The alternative, an `if (is_english) ... else ...` in every HTML string, is how the HilFo report is written, and it is easy to overlook one of them.

```{r p4_labels}
bilingual_report <- function(responses, item_bank,
                             demographics = NULL, rv = NULL) {
  is_english <- is.null(rv) || identical(shiny::isolate(rv$language), "en")

  L <- if (is_english) {
    list(
      title  = "Your Personality Profile",
      note   = "Scale: 1 = strongly disagree, 5 = strongly agree",
      thanks = "Thank you for participating!",
      labels = c(Extraversion      = "Extraversion",
                 Agreeableness     = "Agreeableness",
                 Conscientiousness = "Conscientiousness",
                 Neuroticism       = "Neuroticism",
                 Openness          = "Openness")
    )
  } else {
    list(
      title  = "Ihr Persönlichkeitsprofil",
      note   = "Skala: 1 = stimme gar nicht zu, 5 = stimme voll zu",
      thanks = "Vielen Dank für Ihre Teilnahme!",
      labels = c(Extraversion      = "Extraversion",
                 Agreeableness     = "Verträglichkeit",
                 Conscientiousness = "Gewissenhaftigkeit",
                 Neuroticism       = "Neurotizismus",
                 Openness          = "Offenheit")
    )
  }

  n <- nrow(item_bank)
  if (length(responses) < n)
    responses <- c(responses, rep(NA, n - length(responses)))
  responses <- as.numeric(responses)

  dims <- names(L$labels)
  sc   <- sapply(dims, function(d)
    score_scale(item_bank$id[item_bank$domain == d], responses, item_bank))

  rows <- vapply(dims, function(d) {
    s <- if (is.na(sc[d])) "&mdash;" else sprintf("%.2f", sc[d])
    sprintf(
      "<tr>
         <td style='padding:10px;border-bottom:1px solid #eee;'>%s</td>
         <td style='padding:10px;text-align:center;font-weight:bold;'>%s</td>
       </tr>", L$labels[[d]], s)
  }, character(1))

  shiny::HTML(paste0(
    "<div style='padding:24px;max-width:700px;margin:0 auto;font-family:sans-serif;'>",
    "<h2 style='text-align:center;'>", L$title, "</h2>",
    "<p style='text-align:center;color:#888;font-size:.9em;'>", L$note, "</p>",
    "<table style='width:100%;border-collapse:collapse;'>",
    paste(rows, collapse = ""), "</table>",
    "<p style='text-align:center;margin-top:20px;color:#4A90D9;'>", L$thanks, "</p>",
    "</div>"
  ))
}
```

```{r p4_bilingual_preview, results = "asis", echo = FALSE}
cat('<div style="border:2px solid #27AE60;border-radius:6px;margin:20px 0;overflow:hidden;">')
cat('<div style="background:#27AE60;color:white;padding:6px 16px;font-size:12px;font-weight:bold;">PARTICIPANT VIEW: bilingual report, English</div>')
show_html(bilingual_report(responses, bfi_items, demographics, NULL))
cat('</div>')
```

------------------------------------------------------------------------

# Part 5: The csv_processor

A report can offer participants their own data as a CSV file. inrep builds this file when the page sends `download_csv_trigger`, usually from a button in the report HTML (the programming anxiety case study contains one):

```html
<button onclick="Shiny.setInputValue('download_csv_trigger', Math.random(), {priority: 'event'})">
  Download my data
</button>
```

The file contains one row with the session information, the demographic answers and the item responses under their IDs (`BFI_01 = 4`, `BFI_02 = 2`, ...). A `csv_processor` receives this row before it is written and can add columns, for example the participant's scale scores.

The `csv_processor` does not change the data inrep stores for the researcher. If the stored data should contain scale scores as well, the simplest way is to build the record yourself in the `completion_handler` of the last page before the report, as HilFo does with `build_hilfo_record()` and `score_hilfo_scales()`. The record then has a fixed set of columns, already scored, and can be saved wherever the study keeps its data.

## 5.1 What the function receives

```{r p5_sig, eval = FALSE}
my_csv <- function(csv_data, responses, demographics, item_bank) {
  # csv_data is the one-row data frame described above; add columns and return it
  csv_data
}
```

## 5.2 Example: adding BFI scale scores

```{r p5_csv_fn}
bfi_csv_processor <- function(csv_data, responses, demographics, item_bank) {
  if (is.null(responses) || length(responses) == 0) return(csv_data)
  n <- nrow(item_bank)
  if (length(responses) < n)
    responses <- c(responses, rep(NA, n - length(responses)))
  responses <- as.numeric(responses)

  dims <- c("Extraversion", "Agreeableness", "Conscientiousness",
            "Neuroticism",  "Openness")
  for (d in dims) {
    ids <- item_bank$id[item_bank$domain == d]
    csv_data[[paste0("BFI_", d)]] <- score_scale(ids, responses, item_bank)
  }

  # slider from 0-100 to 1-5, as in Part 2
  sl <- suppressWarnings(as.numeric(demographics$stress_level))
  csv_data$stress_1to5 <- if (!is.na(sl)) round(sl / 100 * 4 + 1, 2) else NA_real_

  csv_data
}
```

The columns added for the simulated participant:

```{r p5_check}
dummy_row  <- data.frame(participant_id = "demo", stringsAsFactors = FALSE)
result_row <- bfi_csv_processor(dummy_row, responses, demographics, bfi_items)
result_row[, grep("^BFI_|stress", names(result_row)), drop = FALSE]
```

## 5.3 Attaching both processors

```{r p5_attach, eval = FALSE}
config <- inrep::create_study_config(
  name              = "BFI study",
  model             = "GRM",
  max_items         = 10,
  results_processor = bilingual_report,   # what the participant sees
  csv_processor     = bfi_csv_processor   # what the researcher gets
)
```

Both processors call the same `score_scale()`. If the keying of an item changes, it changes in one place, and the report and the download cannot disagree.

------------------------------------------------------------------------

# Part 6: Comparing against norms

A mean of 3.4 on a 1-5 scale is hard to interpret without a reference. Participants usually understand a comparison more easily: a score of 3.4 that is higher than that of 68% of the other participants. There are two sources for such a reference.

| Reference | Available from the first participant | Changes during the study |
|---|---|---|
| a fixed norm sample | yes | no |
| the data collected so far | only after enough participants | yes |

## 6.1 A fixed norm sample

The reference scores are prepared once and shipped with the study. In this vignette `norm_pop` is simulated in the setup chunk; in a real study it would be a published norm sample or the data of an earlier cohort. `pct_rank()` returns the percentage of reference scores below the participant's score. For the simulated participant:

```{r p6_norms}
percentiles <- sapply(names(norm_pop),
  function(d) pct_rank(scores[[d]], norm_pop[[d]]))
percentiles
```

Part 7 combines `pct_rank()` with the radar chart and the norm curve.

## 6.2 The data collected so far

If the study saves one file per participant, as HilFo does in `study_data/hilfo_results/`, the report can compare each new participant with everyone before them by reading that folder when the report is built. The folder and the score columns are those of your own record; inrep does not write such a folder by itself.

```{r p6_live_loader}
load_prior_data <- function(data_folder, score_cols) {
  if (is.null(data_folder) || !dir.exists(data_folder)) return(NULL)
  files <- list.files(data_folder, pattern = "[.]csv$", full.names = TRUE)
  if (length(files) == 0) return(NULL)
  rows <- lapply(files, function(f)
    tryCatch(read.csv(f)[, intersect(score_cols, names(read.csv(f))), drop = FALSE],
             error = function(e) NULL))
  rows <- Filter(Negate(is.null), rows)
  if (length(rows) == 0) return(NULL)
  do.call(rbind, rows)
}
```

The processor below is created by a function that remembers the folder (a closure). Each time a report is built, it reads the folder as it is at that moment, and until ten participants have been saved it uses the fixed norm sample instead. Ten is an arbitrary minimum; with very few reference scores a percentile mostly reflects who happened to take part first.

```{r p6_live_processor, eval = FALSE}
make_live_processor <- function(data_folder = NULL, norm_fallback = norm_pop) {
  score_cols <- paste0("BFI_",
    c("Extraversion", "Agreeableness", "Conscientiousness", "Neuroticism", "Openness"))

  function(responses, item_bank, demographics = NULL, session = NULL) {
    prior <- load_prior_data(data_folder, score_cols)
    ref   <- if (!is.null(prior) && nrow(prior) >= 10) prior else norm_fallback

    n <- nrow(item_bank)
    if (length(responses) < n)
      responses <- c(responses, rep(NA, n - length(responses)))
    responses <- as.numeric(responses)

    dims <- c("Extraversion", "Agreeableness", "Conscientiousness",
              "Neuroticism",  "Openness")
    sc   <- setNames(sapply(dims, function(d)
      score_scale(item_bank$id[item_bank$domain == d], responses, item_bank)), dims)
    pcts <- setNames(sapply(dims, function(d)
      pct_rank(sc[d], ref[[paste0("BFI_", d)]])), dims)

    note <- if (!is.null(prior) && nrow(prior) >= 10)
      paste0("Compared to ", nrow(ref), " participants who completed this study before you")
    else
      paste0("Compared to a reference population (N = ", nrow(ref), ")")

    # build the table as in Part 3, with pcts as an additional column
    shiny::HTML(paste0("<p>", note, "</p>"))
  }
}

config <- inrep::create_study_config(
  name              = "BFI longitudinal",
  model             = "GRM",
  max_items         = 10,
  results_processor = make_live_processor(data_folder = "study_data/my_results",
                                          norm_fallback = norm_pop),
  csv_processor     = bfi_csv_processor
)
```

------------------------------------------------------------------------

# Part 7: A complete example

The script below puts the parts together: a short Big Five study in English with demographics, two item pages, the opt-out from Part 4, and a report with radar chart, table with percentiles and a norm curve. It expects `svg_radar()` and `svg_norm_curve()` from Part 3 in a file `report_helpers.R` next to the script.

```{r p7_full, eval = FALSE}
# Big Five study with a personal report
library(inrep)
library(shiny)
source("report_helpers.R")   # svg_radar(), svg_norm_curve() from Part 3

# Item bank ----
data(bfi_items)
bfi_items$id <- paste0("BFI_", sprintf("%02d", seq_len(nrow(bfi_items))))
bfi_domains <- list(
  Extraversion      = c(1, 2, 3, 14, 18, 22, 25),
  Agreeableness     = c(4, 5, 11, 12, 13, 19, 26),
  Conscientiousness = c(6, 17, 24, 27, 30),
  Neuroticism       = c(7, 8, 16, 21, 23, 28),
  Openness          = c(9, 10, 15, 20, 29)
)
bfi_items$domain <- NA_character_
for (d in names(bfi_domains)) bfi_items$domain[bfi_domains[[d]]] <- d
bfi_items$reverse_coded <- seq_len(nrow(bfi_items)) %in% c(2, 5, 8, 12, 14, 21, 28)

# Scoring, used by both processors ----
score_scale <- function(ids, responses, item_bank, max_response = 5) {
  idx <- match(ids, item_bank$id)
  raw <- responses[idx]
  rev <- item_bank$reverse_coded[idx]
  raw[rev] <- (max_response + 1) - raw[rev]
  if (sum(!is.na(raw)) < 2) NA_real_ else mean(raw, na.rm = TRUE)
}
safe_sd <- function(x) { v <- x[!is.na(x)]; if (length(v) < 2) NA_real_ else sd(v) }
pct_rank <- function(score, reference) {
  if (is.na(score) || length(reference) < 5) return(NA_real_)
  round(mean(reference < score, na.rm = TRUE) * 100)
}

# Reference sample (simulated, N = 300) ----
set.seed(100)
norm_pop <- data.frame(
  Extraversion      = pmin(5, pmax(1, rnorm(300, 3.2, 0.7))),
  Agreeableness     = pmin(5, pmax(1, rnorm(300, 3.7, 0.6))),
  Conscientiousness = pmin(5, pmax(1, rnorm(300, 3.5, 0.6))),
  Neuroticism       = pmin(5, pmax(1, rnorm(300, 2.9, 0.8))),
  Openness          = pmin(5, pmax(1, rnorm(300, 3.4, 0.6)))
)

# Demographics ----
demographic_configs <- list(
  age = list(
    question    = "How old are you?", question_en = "How old are you?",
    type = "numeric", required = TRUE
  ),
  gender = list(
    question    = "What is your gender?", question_en = "What is your gender?",
    type = "select",
    options    = c("Female" = "f", "Male" = "m", "Diverse" = "d"),
    options_en = c("Female" = "f", "Male" = "m", "Diverse" = "d"),
    required   = TRUE
  ),
  show_personal_results = list(
    question    = "Would you like to see your personality profile at the end?",
    question_en = "Would you like to see your personality profile at the end?",
    type = "select",
    options    = c("Yes" = "yes", "No" = "no"),
    options_en = c("Yes" = "yes", "No"  = "no"),
    required   = TRUE
  )
)

# Page flow ----
# all five domains on two pages, so that the radar has five axes
idx_1 <- which(bfi_items$domain %in% c("Extraversion", "Agreeableness", "Conscientiousness"))
idx_2 <- which(bfi_items$domain %in% c("Neuroticism", "Openness"))

custom_page_flow <- list(
  list(id = "intro",  type = "custom",
       title = "Welcome", title_en = "Welcome",
       content    = "<h2>Welcome</h2><p>This takes about 5 minutes.</p>",
       content_en = "<h2>Welcome</h2><p>This takes about 5 minutes.</p>"),
  list(id = "demos",  type = "demographics",
       title = "About you", title_en = "About you",
       demographics = names(demographic_configs), required = TRUE),
  list(id = "block_1", type = "items",
       title = "Block 1", title_en = "Block 1",
       item_indices    = idx_1, scale_type = "likert",
       instructions    = "Rate how well each statement describes you.",
       instructions_en = "Rate how well each statement describes you."),
  list(id = "block_2", type = "items",
       title = "Block 2", title_en = "Block 2",
       item_indices    = idx_2, scale_type = "likert",
       instructions    = "Rate how well each statement describes you.",
       instructions_en = "Rate how well each statement describes you."),
  list(id = "results", type = "results",
       title = "Your profile", title_en = "Your profile")
)

# Study configuration ----
study_config <- inrep::create_study_config(
  name                = "Big Five Study",
  custom_page_flow    = custom_page_flow,
  demographics        = names(demographic_configs),
  demographic_configs = demographic_configs,
  model               = "GRM",
  adaptive            = FALSE,
  min_items           = nrow(bfi_items),
  max_items           = nrow(bfi_items),
  language            = "en",
  bilingual           = FALSE,
  response_ui_type    = "radio",
  progress_style      = "bar",
  session_save        = TRUE
)

# Report ----
#
# Three parts: the radar shows the shape of the profile, the table the exact
# scores with their percentile in the reference sample, and the norm curve
# where the highest score lies in that sample.
study_config$results_processor <- function(responses, item_bank,
                                           demographics = NULL) {
  tryCatch({
    # opt-out first (Part 4)
    opted_out <- isTRUE(
      tolower(as.character(demographics$show_personal_results)) %in%
        c("no", "n", "false", "0"))
    if (opted_out) {
      return(shiny::HTML(
        "<div style='padding:40px;text-align:center;font-family:sans-serif;'>
           <h2>Thank you. Your data has been saved.</h2></div>"
      ))
    }

    n <- nrow(item_bank)
    if (length(responses) < n)
      responses <- c(responses, rep(NA, n - length(responses)))
    responses <- as.numeric(responses)
    n_done    <- sum(!is.na(responses))

    dims <- c("Extraversion", "Agreeableness", "Conscientiousness",
              "Neuroticism",  "Openness")
    sc   <- setNames(sapply(dims, function(d)
      score_scale(item_bank$id[item_bank$domain == d], responses, item_bank)), dims)
    sds  <- setNames(sapply(dims, function(d) {
      idx <- match(item_bank$id[item_bank$domain == d], item_bank$id)
      safe_sd(responses[idx])
    }), dims)
    pcts <- setNames(sapply(dims, function(d)
      pct_rank(sc[d], norm_pop[[d]])), dims)

    # radar
    radar_html <- paste0(
      "<h3 style='text-align:center;font-family:system-ui,sans-serif;",
      "color:#2c3e50;margin:24px 0 4px;'>Profile</h3>",
      "<p style='text-align:center;font-family:system-ui,sans-serif;",
      "font-size:.85em;color:#9ca3af;margin:0 0 4px;'>",
      "Profile shape across all five dimensions (scale 1-5)</p>",
      svg_radar(sc)
    )

    # table with percentiles
    rows <- vapply(dims, function(d) {
      pct    <- if (!is.na(pcts[d])) pcts[d] else 0
      bar_w  <- max(0, min(100, pct))
      m_str  <- if (is.na(sc[d]))   "-" else sprintf("%.2f", sc[d])
      sd_str <- if (is.na(sds[d]))  "-" else sprintf("%.2f", sds[d])
      p_str  <- if (is.na(pcts[d])) "?" else paste0(pcts[d], "%")
      sprintf(
        "<tr style='border-bottom:1px solid #f3f4f6;'>
           <td style='padding:10px 14px;font-weight:500;font-family:system-ui;'>%s</td>
           <td style='padding:10px 14px;text-align:center;font-weight:600;color:#4A90D9;
                      font-family:system-ui;'>%s</td>
           <td style='padding:10px 14px;text-align:center;color:#6b7280;
                      font-family:system-ui;'>%s</td>
           <td style='padding:10px 14px;text-align:center;color:#27AE60;
                      font-weight:600;font-family:system-ui;'>%s</td>
           <td style='padding:10px 14px;'>
             <div style='background:#e5e7eb;border-radius:4px;height:7px;'>
               <div style='background:#27AE60;height:100%%;border-radius:4px;width:%d%%;'></div>
             </div>
           </td>
         </tr>", d, m_str, sd_str, p_str, bar_w)
    }, character(1))

    table_html <- paste0(
      "<h3 style='text-align:center;font-family:system-ui,sans-serif;",
      "color:#2c3e50;margin:24px 0 4px;'>Scores &amp; Comparison</h3>",
      "<p style='text-align:center;font-family:system-ui,sans-serif;",
      "font-size:.85em;color:#9ca3af;margin:0 0 8px;'>",
      "Compared to a reference population of 300 participants</p>",
      "<div style='border:1px solid #e5e7eb;border-radius:8px;overflow:hidden;",
      "max-width:700px;margin:0 auto;'>",
      "<table style='width:100%;border-collapse:collapse;'>",
      "<thead><tr style='background:#f9fafb;border-bottom:2px solid #e5e7eb;'>",
      "<th style='padding:12px 14px;text-align:left;font-family:system-ui;'>Scale</th>",
      "<th style='padding:12px 14px;text-align:center;font-family:system-ui;'>M</th>",
      "<th style='padding:12px 14px;text-align:center;font-family:system-ui;'>SD</th>",
      "<th style='padding:12px 14px;text-align:center;font-family:system-ui;'>Pct.</th>",
      "<th style='padding:12px 14px;width:25%;font-family:system-ui;'></th>",
      "</tr></thead><tbody>", paste(rows, collapse = ""), "</tbody></table>",
      "<div style='padding:8px 14px;background:#f9fafb;border-top:1px solid #e5e7eb;",
      "font-size:0.75rem;color:#9ca3af;font-family:system-ui;'><em>",
      "M = Mean. SD = Standard deviation. Scale: 1-5. ",
      "Pct. = percentile relative to reference population (N = 300).",
      "</em></div></div>"
    )

    # norm curve for the highest score
    top_dim  <- dims[which.max(sc)]
    curve_html <- ""
    if (!is.na(sc[top_dim]) && top_dim %in% names(norm_pop)) {
      ref_vec   <- norm_pop[[top_dim]]
      curve_html <- paste0(
        "<h3 style='text-align:center;font-family:system-ui,sans-serif;",
        "color:#2c3e50;margin:24px 0 4px;'>Where You Fall: ", top_dim, "</h3>",
        "<p style='text-align:center;font-family:system-ui,sans-serif;",
        "font-size:.85em;color:#9ca3af;margin:0 0 4px;'>",
        "Your score on the dimension where you scored highest, ",
        "relative to the reference sample</p>",
        svg_norm_curve(
          score        = sc[top_dim],
          norm_mean    = mean(ref_vec, na.rm = TRUE),
          norm_sd      = sd(ref_vec,   na.rm = TRUE),
          label        = paste0("Reference population (N = 300)"),
          color        = "#8E44AD",
          shade_color  = "rgba(142,68,173,0.18)"
        )
      )
    }

    shiny::HTML(paste0(
      "<div style='max-width:760px;margin:0 auto;padding:24px 16px;'>",
      "<h2 style='text-align:center;color:#2c3e50;margin:0 0 4px 0;",
      "font-family:system-ui,sans-serif;font-weight:700;'>Your Big Five Profile</h2>",
      "<p style='text-align:center;font-family:system-ui,sans-serif;",
      "font-size:.85em;color:#9ca3af;margin:0 0 16px 0;'>Based on ",
      n_done, " items</p>",
      radar_html,
      table_html,
      curve_html,
      "</div>"
    ))

  }, error = function(e) {
    shiny::HTML(sprintf(
      "<div style='padding:20px;color:#900;'>
         <h3>Report could not be generated</h3>
         <p>%s</p></div>", e$message))
  })
}

# Scale scores in the CSV participants can download ----
study_config$csv_processor <- function(csv_data, responses, demographics, item_bank) {
  if (is.null(responses) || length(responses) == 0) return(csv_data)
  n <- nrow(item_bank)
  if (length(responses) < n)
    responses <- c(responses, rep(NA, n - length(responses)))
  responses <- as.numeric(responses)

  dims <- c("Extraversion", "Agreeableness", "Conscientiousness",
            "Neuroticism",  "Openness")
  for (d in dims)
    csv_data[[paste0("BFI_", d)]] <-
      score_scale(item_bank$id[item_bank$domain == d], responses, item_bank)
  csv_data
}

# Start ----
inrep::launch_study(
  config     = study_config,
  item_bank  = bfi_items,
  debug_mode = FALSE
)
```

------------------------------------------------------------------------

# Common mistakes

Most problems with reports we have seen come from a few recurring mistakes.

1. **Returning a character string.** The processor has to return `shiny::HTML(...)`; a plain string is displayed as text, tags included.

2. **Not padding `responses`.** In adaptive studies and after skipped pages the vector is shorter than the item bank. Without padding, positions shift and scores are computed from the wrong items without any error.

3. **Scoring by position.** `responses[1:6]` is only correct as long as nobody reorders the item bank. Looking items up with `match(ids, item_bank$id)` stays correct.

4. **Assuming the item order of a built-in bank.** `bfi_items` is not sorted by domain; its keys are set in the setup chunk of this vignette.

5. **Looking for sliders in `responses`.** Slider values are in `demographics`; `responses[30]` for a slider is simply `NA`.

6. **Reading the language from `session$input`.** On the final results page `session$input` has no `language` element. The participant's language is in `rv$language`.

7. **Scoring before the opt-out check.** Check `show_personal_results` first. Participants who declined the report should not wait for charts they will not see.

8. **Writing `if (is_english)` into every string.** One list of labels per language at the top of the processor is easier to keep complete.

9. **Expecting the `csv_processor` to change the stored data.** It only shapes the CSV participants download. Scores for the researcher belong in the record the study saves (Part 5).

10. **Expecting different content on two results pages from one processor.** The study-level processor runs on every results page. A page that should show something else needs its own `results_processor` (Part 1.4).
