---
title: "How to Calibrate Your Item Bank with inrep"
author: "inrep"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: cerulean
vignette: >
  %\VignetteIndexEntry{How to Calibrate Your Item Bank with inrep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = FALSE
)
```

## Overview

**Item calibration** is the process of estimating IRT model parameters -- item location
(difficulty), discrimination, and for polytomous items the threshold boundaries -- from
empirical response data. Calibrated parameters are what transform a list of questions
into a proper item bank that an adaptive assessment engine can use.

This vignette walks through a full calibration workflow using the **Big Five Inventory
(BFI)** from the `psych` package as the target questionnaire. The BFI uses a 6-point
Likert scale and measures five personality dimensions (Agreeableness, Conscientiousness,
Extraversion, Neuroticism, Openness). It is publicly available and well-validated,
making it ideal for illustrating calibration from start to finish.

`inrep` also ships a built-in dataset, `bfi_items`, but that object serves a different
purpose: it is a **simulated, pre-calibrated 30-item demonstration bank** with 5 response
categories. This vignette instead starts from the raw 25-item `psych::bfi` responses and
shows how a calibrated bank is produced in the first place.

**Scenario**: We calibrate item parameters from the 2,436-person BFI dataset in
`psych::bfi`. Those calibrated parameters are the prior that inrep uses when a new
individual completes the deployed study -- each participant is scored against the
pre-calibrated bank in real time and receives a results page. All code is fully
reproducible.

> **Note on running this vignette.** All code chunks have `eval = FALSE` because TAM
> model fitting takes several minutes to run. Expected outputs are shown as inline
> comments. To reproduce everything yourself, copy each chunk into a fresh R session
> and run them in order.

---

## 1. Prerequisites

```{r packages}
# Install these packages first if you don't have them yet:
# install.packages(c("psych", "TAM"))
# devtools::install_github("selvastics/inrep")
library(psych)   # source of BFI item data
library(TAM)     # IRT model fitting and diagnostics
library(inrep)   # adaptive assessment framework
```

---

## 2. Load and Inspect the Data

```{r load_data}
data(bfi)

# BFI items are columns 1-25 (A1-A5, C1-C5, E1-E5, N1-N5, O1-O5)
# Columns 26-28: gender, education, age (not used in calibration)
items_raw <- bfi[, 1:25]

cat("Dimensions:", nrow(items_raw), "respondents x", ncol(items_raw), "items\n")
cat("Scale range: 1 to", max(items_raw, na.rm = TRUE), "(6-point Likert)\n")
cat("Complete cases:", sum(complete.cases(items_raw)), "\n")

psych::describe(items_raw)[, c("n", "mean", "sd", "skew", "kurtosis")]
```

---

## 3. Reverse Coding

Seven BFI items are reverse-scored: a *low* rating reflects *high* standing on the
trait. We reverse-code **before** calibration and before the dimensionality check.
Mixing polarities in a factor analysis or IRT model produces threshold parameters
that are inconsistent in sign and breaks model fit.

```{r reverse_code}
# Reverse-scored items in the psych BFI (6-point scale: recoded as 7 - x)
#   A1  "Am indifferent to the feelings of others"
#   C4  "Do things in a half-way manner"
#   C5  "Waste my time"
#   E1  "Don't talk a lot"
#   E2  "Find it difficult to approach others"
#   O2  "Avoid difficult reading material"
#   O5  "Will not probe deeply into a subject"
reverse_items <- c("A1", "C4", "C5", "E1", "E2", "O2", "O5")

items_rc <- items_raw
for (i in reverse_items) {
  items_rc[[i]] <- 7 - items_raw[[i]]
}

# Listwise deletion: keep only complete cases
items_rc <- items_rc[complete.cases(items_rc), ]
cat("N after listwise deletion:", nrow(items_rc), "\n")
# N after listwise deletion: 2436
```

> **Missing data**: TAM's EM algorithm handles within-row missingness under MAR
> via partial likelihoods -- listwise deletion is not required. We use it here
> for pedagogical simplicity. For studies with substantial or non-ignorable
> missingness, consider multiple imputation before calibration.

---

## 4. Dimensionality Check

Before fitting five separate unidimensional models, we confirm the five-factor structure
in the data. Calibrating items together that belong to different dimensions violates
local independence and biases parameter estimates.

```{r dim_check_parallel}
# Parallel analysis to determine the number of factors
# n.iter = 20 for speed; increase to 100 for publication-level work
fa_para <- psych::fa.parallel(
  items_rc, fm = "ml", fa = "fa", n.iter = 20,
  main = "Parallel analysis -- BFI 25 items"
)
# Expected result: eigenvalues for 5 factors exceed the simulated reference line
```

```{r dim_check_omega}
# Define the five BFI subscales
dims <- list(
  Agreeableness     = c("A1", "A2", "A3", "A4", "A5"),
  Conscientiousness = c("C1", "C2", "C3", "C4", "C5"),
  Extraversion      = c("E1", "E2", "E3", "E4", "E5"),
  Neuroticism       = c("N1", "N2", "N3", "N4", "N5"),
  Openness          = c("O1", "O2", "O3", "O4", "O5")
)

# Within-scale omega confirms adequate unidimensionality per dimension
# omega_h >= 0.50 and omega_t >= 0.70 are acceptable thresholds
for (d in names(dims)) {
  om <- psych::omega(items_rc[, dims[[d]]], plot = FALSE)
  cat(sprintf("%-20s  omega_h = %.3f   omega_t = %.3f\n",
              d, om$omega_h, om$omega.tot))
}
```

```
# Expected output (approximate; omega uses simulation so minor variation is normal):
# Agreeableness         omega_h = 0.69   omega_t = 0.76
# Conscientiousness     omega_h = 0.63   omega_t = 0.78
# Extraversion          omega_h = 0.67   omega_t = 0.80
# Neuroticism           omega_h = 0.73   omega_t = 0.85
# Openness              omega_h = 0.56   omega_t = 0.67
```

All five scales clear the omega_h >= 0.50 threshold. We proceed with separate
unidimensional PCMs for each dimension.

---

## 5. Calibrated Parameters as the Prior for Deployment

Once calibration is complete, the item bank parameters are fixed. They serve as the
**prior** for every subsequent individual who completes the study. When a participant
runs the deployed assessment via `launch_study()`, inrep estimates their theta in real
time using `estimate_ability()` and the calibrated bank -- it does not re-fit the IRT
model. The participant then sees a personalized results page.

Periodic re-calibration (re-running `tam.mml()` on a larger or refreshed dataset) is a
maintenance task, not a per-participant operation. When genuine new response data are
available, merge them with the original calibration sample and re-run the workflow from
Section 6 onwards:

```{r recalibrate_workflow, eval=FALSE}
# When genuine new data arrive from a later collection wave:
data_wave2     <- read.csv("wave2_responses.csv")    # real new participants
cal_combined   <- rbind(items_rc, data_wave2)         # genuine additional information
# ... then re-run from Section 6 with cal_combined as input
```

For the remainder of this vignette we use `items_rc` (N = 2,436) as the calibration
sample.

---

## 6. Fitting the Calibration Models

We fit **separate models for each Big Five dimension** because the five dimensions are
theoretically distinct latent traits (confirmed by the dimensionality check in Section 4).
A joint 25-item unidimensional model would violate local independence.

### 6a. Partial Credit Model (PCM, Rasch family) -- recommended for personality scales

`TAM::tam.mml()` with polytomous responses fits the **Partial Credit Model**, the
polytomous generalization of the Rasch model (Masters, 1982) that constrains all item
discrimination parameters to 1. We use it as the starting point for personality
questionnaires because:

- Items are written to be equally discriminating by design.
- Rasch models provide interval-level measurement and are easier to validate.
- Estimation is faster and more stable than GPCM for small-to-moderate samples.

```{r fit_pcm}
models <- list()
for (d in names(dims)) {
  sub <- items_rc[, dims[[d]]]
  sub <- sub[complete.cases(sub), ]   # per-dimension listwise: best practice for real data
                                      # (items_rc is already complete, so this is a no-op
                                      # here; shown for generalizability to workflows
                                      # with partial item-level missingness)
  mod <- TAM::tam.mml(
    resp    = sub,
    verbose = FALSE,
    control = list(maxiter = 500, convD = 0.001)
  )
  models[[d]] <- mod
  cat(d, "| log-likelihood:", round(mod$ic$loglike, 1),
      "| AIC:", round(mod$ic$AIC, 1), "\n")
}
```

```
# Agreeableness     | log-likelihood: -17520.9 | AIC: 35103.9
# Conscientiousness | log-likelihood: -18497.5 | AIC: 37056.9
# Extraversion      | log-likelihood: -19021.3 | AIC: 38104.7
# Neuroticism       | log-likelihood: -19376.4 | AIC: 38814.8
# Openness          | log-likelihood: -17895.8 | AIC: 35853.7
```

---

## 7. Extracting Threshold Parameters

`TAM::tam.threshold()` returns the **boundary thresholds** -- the theta values where
a respondent is equally likely to choose a response below vs. at-or-above that
boundary. For a 6-point BFI scale, TAM internally creates a theoretical response
category 0 (since raw responses range 1-6), which produces **six threshold columns
(Cat1-Cat6)**:

- **Cat1 = -12**: the theoretical threshold between "would respond 0" and "would
  respond 1." Because no BFI respondent ever answers 0, TAM estimates this boundary
  at the extreme lower end (-12). Discard Cat1.
- **Cat2-Cat6 = b1-b5**: the five substantive thresholds for the 1->2, 2->3, 3->4,
  4->5, 5->6 transitions. These map directly to the `b1, b2, b3, b4, b5` columns
  inrep expects.

```{r extract_thresholds}
thresholds <- lapply(models, TAM::tam.threshold)
lapply(thresholds, round, 3)
```

```
# $Agreeableness
#    Cat1   Cat2   Cat3   Cat4   Cat5  Cat6
# A1  -12 -2.140 -1.215 -0.734 -0.374 0.478
# A2  -12 -2.281 -1.470 -1.173 -0.581 0.622
# A3  -12 -1.889 -1.230 -0.948 -0.408 0.787
# A4  -12 -1.740 -1.042 -0.817 -0.421 0.202
# A5  -12 -2.287 -1.318 -0.928 -0.313 0.891
#
# $Conscientiousness
#    Cat1   Cat2   Cat3   Cat4   Cat5  Cat6
# C1  -12 -2.199 -1.424 -0.974 -0.309 1.055
# C2  -12 -2.185 -1.207 -0.801 -0.171 1.134
# C3  -12 -2.218 -1.213 -0.810 -0.047 1.282
# C4  -12 -2.422 -1.335 -0.659 -0.212 0.705
# C5  -12 -1.509 -0.607 -0.024  0.254 1.033
#
# $Extraversion
#    Cat1   Cat2   Cat3   Cat4   Cat5  Cat6
# E1  -12 -1.551 -0.766 -0.302  0.030 0.815
# E2  -12 -1.505 -0.738 -0.150  0.146 1.036
# E3  -12 -1.847 -1.046 -0.551  0.310 1.548
# E4  -12 -1.808 -1.018 -0.696 -0.298 0.833
# E5  -12 -2.072 -1.195 -0.804 -0.202 1.074
#
# $Neuroticism
#    Cat1   Cat2   Cat3   Cat4  Cat5  Cat6
# N1  -12 -0.911 -0.054  0.347 0.972 1.863
# N2  -12 -1.536 -0.547 -0.128 0.639 1.652
# N3  -12 -1.201 -0.232  0.097 0.762 1.767
# N4  -12 -1.263 -0.251  0.152 0.841 1.679
# N5  -12 -0.883 -0.033  0.314 0.901 1.672
#
# $Openness
#    Cat1   Cat2   Cat3   Cat4   Cat5  Cat6
# O1  -12 -2.531 -1.536 -1.102 -0.459 0.536
# O2  -12 -1.493 -0.852 -0.428 -0.150 0.552
# O3  -12 -1.885 -1.266 -0.869 -0.130 1.077
# O4  -12 -2.012 -1.288 -1.033 -0.590 0.282
# O5  -12 -1.961 -1.195 -0.698 -0.258 0.683
```

**Reading the thresholds**: For A1 (reverse-coded), b1 = -2.133 means respondents
with theta < -2.133 are more likely to choose response category 1. At b5 = 0.462,
P(X >= 6) = 0.5 -- respondents at this theta have a 50% chance of choosing category
6 vs. any lower category; above b5, category 6 becomes the most likely response.
Neuroticism thresholds are higher (nearer to 0) because most general-population
respondents score in the low-to-moderate range on neuroticism items.

---

## 8. Evaluating Item Fit

We do not skip item fit -- even if TAM converges cleanly, individual items may misfit.

```{r item_fit}
# Compute fit statistics for each dimension
fit_stats <- lapply(models, TAM::tam.fit)

# Infit MNSQ -- primary diagnostic (information-weighted):
#   Acceptable range: 0.6 - 1.4
#   > 1.4 : unpredictable / noisy responses
#   < 0.6 : overly predictable / item redundancy
#
# Outfit MNSQ -- unweighted; more sensitive to unexpected extreme responses

# Aggregate to item level for Agreeableness
fi_A <- fit_stats$Agreeableness$itemfit
fi_A$item <- sub("_Cat.*", "", fi_A$parameter)
item_fit_A <- aggregate(cbind(Infit, Outfit) ~ item, data = fi_A, FUN = mean)
print(round(item_fit_A, 3))
```

```
#   item  Infit Outfit
# 1   A1  0.945  2.928   <- Outfit far outside [0.6, 1.4]; Infit within range
# 2   A2  0.788  1.214
# 3   A3  0.777  0.780
# 4   A4  0.898  1.330
# 5   A5  0.829  1.436
```

> **A1**: Infit = 0.95 is within the acceptable range; Outfit = 2.93 far exceeds
> the upper bound of 1.4. Outfit is more sensitive to unexpected responses at the
> extremes of the trait continuum -- here driven by the very low frequency of
> "1" responses on A1 (a floor effect). In the function below we flag using
> **Infit** as the primary criterion, which is standard practice for Rasch-family
> models. If Outfit elevation is a concern, inspect the category frequency
> distribution and consider collapsing the lowest two response categories.

```{r flag_misfitting}
flag_misfitting <- function(fit_obj, lo = 0.6, hi = 1.4) {
  fi <- fit_obj$itemfit
  fi$item <- sub("_Cat.*", "", fi$parameter)
  agg     <- aggregate(Infit ~ item, data = fi, FUN = mean)
  flagged <- agg$item[agg$Infit > hi | agg$Infit < lo]
  if (length(flagged) == 0) cat("  All items within Infit range [0.6, 1.4].\n")
  else cat("  Infit-flagged:", paste(flagged, collapse = ", "), "\n")
}

for (d in names(fit_stats)) {
  cat(d, ":\n"); flag_misfitting(fit_stats[[d]])
}
```

```
# Agreeableness     : All items within Infit range [0.6, 1.4].
# Conscientiousness : All items within Infit range [0.6, 1.4].
# Extraversion      : All items within Infit range [0.6, 1.4].
# Neuroticism       : All items within Infit range [0.6, 1.4].
# Openness          : All items within Infit range [0.6, 1.4].
```

All 25 BFI items pass the Infit criterion. The parameters from Section 7 are trustworthy
for use in an adaptive item bank.

---

## 9. Calibration for Binary Items: 1PL and 2PL

When the item bank contains **binary (correct/incorrect) items** -- knowledge tests,
diagnostic screeners -- we use a different model family.

### 9a. 1PL (Rasch) for binary items

```{r binary_1pl}
# Simulate binary knowledge items with varying discrimination (0/1, correct/incorrect)
# We use a 2PL data-generating model so that anova(mod_1pl, mod_2pl) shows a real
# likelihood-ratio gain when we compare models in 9b.
set.seed(7)
n_items <- 20; n_resp <- 500
slopes  <- runif(n_items, 0.5, 2.0)   # varying discrimination
diff    <- runif(n_items, -2, 2)       # varying difficulty
theta   <- rnorm(n_resp)
prob    <- sapply(seq_len(n_items), function(j) plogis(slopes[j] * (theta - diff[j])))
binary_data <- matrix(
  rbinom(n_resp * n_items, 1, prob),
  nrow = n_resp, ncol = n_items,
  dimnames = list(NULL, paste0("item_", seq_len(n_items)))
)

# Fit 1PL (Rasch): discrimination constrained to 1 for all items
mod_1pl <- TAM::tam.mml(resp = binary_data, verbose = FALSE)

# Item locations (b parameters)
b_1pl <- mod_1pl$item[, "xsi.item"]
round(b_1pl, 3)
```

### 9b. 2PL for binary items

When items are expected to differ in discriminating power (mixed content, varying
cognitive demand), we use 2PL, which adds a per-item discrimination parameter `a`.
Because the data above were generated with varying slopes, `anova(mod_1pl, mod_2pl)`
will show a real likelihood-ratio gain here.

```{r binary_2pl}
mod_2pl <- TAM::tam.mml.2pl(resp = binary_data, verbose = FALSE)

# Difficulty parameters
b_2pl <- mod_2pl$item[, "xsi.item"]

# Discrimination (slope) parameters
# For binary 2PL in TAM >= 4.x, slopes are stored in the column "B.Cat1.Dim1"
# Use grep() as a safeguard if column names change across minor versions:
a_col <- grep("^B\\.Cat1", names(mod_2pl$item), value = TRUE)[1]
a_2pl <- mod_2pl$item[, a_col]

round(data.frame(a = a_2pl, b = b_2pl), 3)
```

```{r compare_models}
# Compare 1PL and 2PL via AIC/BIC and likelihood-ratio test
anova(mod_1pl, mod_2pl)
```

> **When to prefer 2PL**: Use it when items plausibly differ in discriminating
> power -- cognitive tests, diagnostic screeners, or mixed-format banks. For
> personality questionnaires written to a consistent style, PCM/1PL is more
> principled and parsimonious.

---

## 10. Building the inrep Item Bank

We assemble the item bank **programmatically from calibration output** rather than
hardcoding threshold values -- this ensures the bank always reflects the most recent
calibration run.

```{r build_item_bank}
item_questions <- c(
  A1 = "Am indifferent to the feelings of others.",   # reverse-scored
  A2 = "Inquire about others' well-being.",
  A3 = "Know how to comfort others.",
  A4 = "Love children.",
  A5 = "Make people feel at ease.",
  C1 = "Am exacting in my work.",
  C2 = "Continue until everything is perfect.",
  C3 = "Do things according to a plan.",
  C4 = "Do things in a half-way manner.",              # reverse-scored
  C5 = "Waste my time.",                               # reverse-scored
  E1 = "Don't talk a lot.",                            # reverse-scored
  E2 = "Find it difficult to approach others.",        # reverse-scored
  E3 = "Know how to captivate people.",
  E4 = "Make friends easily.",
  E5 = "Take charge.",
  N1 = "Get angry easily.",
  N2 = "Get irritated easily.",
  N3 = "Have frequent mood swings.",
  N4 = "Often feel blue.",
  N5 = "Panic easily.",
  O1 = "Am full of ideas.",
  O2 = "Avoid difficult reading material.",            # reverse-scored
  O3 = "Carry the conversation to a higher level.",
  O4 = "Spend time reflecting on things.",
  O5 = "Will not probe deeply into a subject."         # reverse-scored
)

# Programmatic threshold extraction
# For BFI 1-6 data, tam.threshold() returns 6 columns (Cat1-Cat6).
# Cat1 = theoretical lower-bound (~-12); b1-b5 = Cat2-Cat6.
thresh_mat <- do.call(rbind, lapply(names(dims), function(d) {
  th <- thresholds[[d]]
  # Drop Cat1 (the extreme lower-bound column); keep Cat2-Cat6 = b1-b5
  th[, colnames(th) != "Cat1", drop = FALSE]
}))
colnames(thresh_mat) <- paste0("b", 1:5)

bfi_item_bank <- data.frame(
  id                 = names(item_questions),
  Question           = unname(item_questions),
  domain             = rep(names(dims), each = 5),
  reverse_coded      = names(item_questions) %in% reverse_items,
  ResponseCategories = "1,2,3,4,5,6",
  a                  = 1,   # PCM: discrimination fixed at 1 (Rasch constraint)
  thresh_mat,
  stringsAsFactors   = FALSE
)

# First, run the package's built-in structure validator
validation <- validate_item_bank(bfi_item_bank, model = "GRM")
print(validation$is_valid)
print(validation$messages)

if (!validation$is_valid) {
  stop(paste(validation$messages, collapse = "\n"))
}

# Second, run an explicit monotonicity check.
# validate_item_bank() confirms required columns and model compatibility;
# this additional check confirms threshold ordering within each item.
mono_ok <- apply(bfi_item_bank[, paste0("b", 1:5)], 1, function(r) all(diff(r) > 0))
cat("All items monotone:", all(mono_ok), "\n")
# All items monotone: TRUE
```

> **Monotonicity is a hard requirement.** `validate_item_bank()` checks that the
> bank is structurally compatible with the target model, but it does not verify
> threshold ordering. If b1 > b2 for any item, the GRM
> probability function becomes non-monotone and produces negative item information.
> Always run this check before deploying a bank in inrep.

Do **not** use `inrep::bfi_items` as a parameter benchmark for this bank. The shipped
dataset is a simulated 30-item, 5-category demonstration object, whereas the bank
constructed here is a 25-item, 6-category calibration derived from `psych::bfi`.
The right comparison is conceptual: this vignette shows how a deployable item bank is
created from raw responses before functions like `launch_study()` and `estimate_ability()`
can use it.

---

## 11. Launching an Adaptive Assessment with the Calibrated Bank

### Single-dimension: Extraversion

```{r launch_extraversion}
extra_bank <- bfi_item_bank[bfi_item_bank$domain == "Extraversion", ]

config_extra <- create_study_config(
  name              = "Extraversion Adaptive Assessment",
  model             = "GRM",
  adaptive          = TRUE,
  criteria          = "MI",        # Maximum Information item selection
  min_items         = 3,
  max_items         = 5,
  min_SEM           = 0.30,        # stop when SE(theta) < 0.30
  theme             = "hildesheim",
  language          = "en"
)

# Item selection criteria — what the choices mean
# ------------------------------------------------
# "MI"  (default) — Maximum Information at the point estimate hat_theta.
#   Fast and standard in operational CAT. Works well once SE is small
#   (long tests, 20+ items). Early in a short assessment SE is typically
#   0.5-0.8, so hat_theta is imprecise and MI may optimize for the wrong
#   ability region.
#
# "MEI" — Maximum Expected Information. Integrates I(theta) over the
#   posterior p(theta | hat_theta, SE), selecting items that are informative
#   across the *full plausible range* of true ability. Substantially more
#   accurate than MI for short scales (< 15 items) where SE stays large.
#   For long tests (30+ items) MI and MEI converge; use MI there.
#   Reference: Veerkamp & Berger (1997), J. Educ. Behav. Stat. 22(2).
#
# "WEIGHTED" — Information × content-balance weights. Useful when item
#   groups must be proportionally represented.
#
# "MFI" — Exposure-penalised MI, to avoid overexposure of high-info items.
#
# Rule of thumb: use "MEI" when max_items < 15; use "MI" otherwise.

launch_study(config_extra, extra_bank)

# Note: with max_items equal to the bank size, every item is administered;
# adaptivity here only optimizes administration order. Real adaptive
# deployments truncate well below the bank size -- see the full 25-item
# example below for a more realistic adaptive setup.
```

### Full 25-item non-adaptive assessment

```{r launch_full_bfi}
config_full <- create_study_config(
  name     = "BFI-25 Personality Assessment",
  model    = "GRM",
  adaptive = FALSE,    # fixed sequential order
  max_items = 25,
  theme    = "hildesheim",
  language = "en"
)

launch_study(config_full, bfi_item_bank)
```

---

## 12. Model Selection Reference

| Situation | Response format | TAM function | inrep `model` |
|---|---|---|---|
| Personality / Likert (equal discrimination) | Polytomous | `tam.mml()` | `"GRM"` |
| Ability / knowledge (equal discrimination) | Binary 0/1 | `tam.mml()` | `"1PL"` |
| Ability / knowledge (varying discrimination) | Binary 0/1 | `tam.mml.2pl()` | `"2PL"` |
| Multiple-choice with guessing | Binary 0/1 | `tam.mml.3pl()` | `"3PL"` |

**Note on GRM vs. GPCM**: inrep's item selection engine (`compute_item_info_single()`)
uses the **Samejima GRM** parameterization with boundary thresholds (b1, b2, ...).
TAM's `tam.mml()` for polytomous data fits the **Partial Credit Model** / GPCM with
step parameters. `TAM::tam.threshold()` converts those step parameters to boundary
thresholds, which is why we always use `tam.threshold()` output -- never raw `xsi`
parameters -- when building an inrep item bank for GRM.

---

## 13. Common Pitfalls

| Mistake | Consequence | Fix |
|---|---|---|
| Skipping reverse coding | Threshold signs flip; model fits wrong direction | Apply `7 - x` before `tam.mml()` |
| Skipping dimensionality check | Local independence violated; biased parameters | Run `fa.parallel()` + `omega()` per scale |
| Using invented threshold values | Biased theta estimates; meaningless information curves | Extract from `TAM::tam.threshold()` |
| Skipping `validate_item_bank()` | Structural mismatches reach runtime | Run `validate_item_bank(bank, model = "GRM")` before deployment |
| Fitting one model to all 25 items | Local independence violated | Fit separate models per dimension |
| Ignoring item fit | Misfitting items corrupt ability estimates | Check Infit MNSQ; flag outside [0.6, 1.4] |
| Non-monotone thresholds in bank | Negative item information; adaptive selection breaks | Run monotonicity check per item |
| Deploying on a different population | Parameter drift; SEM underestimated | Re-calibrate or link parameters |
| Reporting plausible values as individual scores | PVs are random posterior draws, not point estimates | Use EAP or WLE for individual reporting (see Section 16) |
| Inflating N by resampling | No new information; overstates precision | Use genuine new participants |

---

## 14. Further Diagnostics and Next Steps

```{r test_information}
# Test information function for the Extraversion bank
mod_E     <- models$Extraversion
theta_seq <- seq(-4, 4, by = 0.1)
info      <- TAM::IRT.informationCurves(mod_E, theta = theta_seq)

plot(theta_seq, rowSums(info$test_info_curves),
     type = "l", lwd = 2, col = "#3498db",
     xlab = expression(theta), ylab = "Test Information",
     main = "Extraversion -- Test Information Function (N=2436, PCM)")
abline(h = 1 / 0.30^2, lty = 2, col = "grey50")   # SEM = 1/sqrt(info) = 0.30
legend("topright",
  c("Test information", "SEM = 0.30 reference"),
       lty = c(1, 2), col = c("#3498db", "grey50"))
```

Additional analyses to consider:

- **DIF analysis**: TAM does not include a dedicated DIF function. Use
  `difR::difMH()` for Mantel-Haenszel DIF on binary items, `lordif::lordif()` for
  ordinal DIF on Likert items, or the separate-group calibration approach in Section 15e.
- **Person fit**: `TAM::tam.personfit()` -- flag aberrant response patterns in
  deployment data.
- **Equating / linking across waves**: `TAM::tam.linking()` fits linking models
  across a list of TAM model objects (see `?TAM::tam.linking`); `sirt::linking.haberman()`
  (Haberman mean-mean / mean-sigma) and the `equateIRT` package are also widely used.
- **Cross-validation**: hold out 20% of your calibration sample, refit, compare
  parameter estimates for stability.

---

## 15. Constraining Parameters During Calibration

Constraints fix one or more IRT parameters to known values (from a prior calibration
study) or force parameters to be equal across items or groups. In TAM, individual
parameter fixing uses `xsi.fixed` (for step/difficulty parameters) or `B.fixed`
(for slope parameters in 2PL models).

> **Clarification on `constraint`**: The `constraint` argument in `tam.mml()` is
> for specifying the *normalization constraint type* (default `"cases"`), not for
> pinning individual parameter values. Use `xsi.fixed` for fixed-value constraints.

### 15a. Fixed-value constraints: anchoring an item to a prior calibration

```{r constraints_identify}
sub_E <- items_rc[, c("E1", "E2", "E3", "E4", "E5")]

# Step 1: fit the unconstrained model to identify xsi parameter indices
mod_E_free <- TAM::tam.mml(resp = sub_E, verbose = FALSE)

# For BFI 1-6 data, TAM creates 6 xsi parameters per item (Cat1-Cat6).
# Row names follow the pattern "itemname_CatK":
cat("xsi parameter names for E1 (first 6 rows):\n")
print(rownames(mod_E_free$xsi)[grepl("^E1_", rownames(mod_E_free$xsi))])
```

```
# [1] "E1_Cat1" "E1_Cat2" "E1_Cat3" "E1_Cat4" "E1_Cat5" "E1_Cat6"
```

```{r constraints_anchor}
# Step 2: fix E1's xsi step parameters to values from a prior calibration.
# This holds E1 as an anchor item so the theta scale does not drift across waves.
#
# IMPORTANT: xsi.fixed pins PCM *step parameters* (the delta's), NOT the
# Thurstonian boundary thresholds returned by tam.threshold(). For polytomous
# items these are different quantities -- fixing threshold values as if they were
# steps produces the wrong response surface. Always source anchor values from
# mod_prior$xsi[..., "xsi"], not from tam.threshold() output.
#
# Here we use mod_E_free's own xsi as a stand-in for a "prior calibration" to
# illustrate the syntax; in practice substitute your previously saved xsi values.
e1_idx_b   <- which(rownames(mod_E_free$xsi) %in% paste0("E1_Cat", 2:6))
e1_xsi_ref <- mod_E_free$xsi[paste0("E1_Cat", 2:6), "xsi"]   # step parameters

mod_E_anchored <- TAM::tam.mml(
  resp      = sub_E,
  xsi.fixed = cbind(e1_idx_b, e1_xsi_ref),
  verbose   = FALSE
)
# E1 parameters are now fixed; E2-E5 are estimated freely
round(mod_E_anchored$xsi[grepl("^E1_", rownames(mod_E_anchored$xsi)), ], 3)
```

> **When to use anchor constraints**: re-calibration with new data, linking across
> collection waves, and confirmatory studies where item parameters are already
> published. Always verify that anchored items still show acceptable fit in the
> new sample before trusting the remaining free parameters.

### 15b. Fixing a slope in a binary 2PL model

For 2PL models, discrimination parameters live in the B matrix and are fixed via
`B.fixed`.

```{r constraints_b_fixed}
# B.fixed: 4-column matrix -- item, category, dimension, value
# For unidimensional binary 2PL:
#   column 1 = item index
#   column 2 = category (2 for Cat1, the discrimination column)
#   column 3 = dimension (1 for unidimensional models)
#   column 4 = fixed value
# To fix item 1's discrimination to 1.5:
B_fixed_mat <- matrix(c(1, 2, 1, 1.5), nrow = 1)   # item, cat, dim, value

mod_2pl_anchored <- TAM::tam.mml.2pl(
  resp    = binary_data,
  B.fixed = B_fixed_mat,
  verbose = FALSE
)
# Check: item 1 discrimination should now be ~1.5
round(mod_2pl_anchored$item[1, grep("^B\\.Cat1", names(mod_2pl_anchored$item))], 3)
```

### 15c. Rating Scale Model: testing equal step structure

The **Rating Scale Model** (RSM, `irtmodel = "RSM"`) is a special case of PCM where
all items share the same set of step difficulties. It is more restrictive than PCM
but useful as a parsimony check.

```{r constraints_rsm}
# Valid TAM irtmodel values: "1PL" (default PCM), "GPCM", "RSM", "PCM2"
mod_rsm <- TAM::tam.mml(
  resp     = sub_E,
  irtmodel = "RSM",
  verbose  = FALSE
)

# Likelihood-ratio test: PCM (more general) vs RSM (constrained)
anova(mod_rsm, mod_E_free)
# Significant chi-squared => PCM is better fit; retain per-item thresholds
```

### 15d. GPCM: freely estimated discrimination per item

If item discriminations are expected to vary (heterogeneous item pool), fit the
Generalized Partial Credit Model.

```{r constraints_gpcm}
mod_gpcm <- TAM::tam.mml(
  resp     = sub_E,
  irtmodel = "GPCM",
  verbose  = FALSE
)

# Compare PCM (a = 1 fixed) vs GPCM (a freely estimated per item)
anova(models$Extraversion, mod_gpcm)
# For BFI items PCM typically fits as well as GPCM, justifying the simpler model
```

### 15e. Simple DIF detection via separate-group calibration

TAM does not include a built-in DIF function. The most direct approach for
polytomous items is separate-group calibration plus parameter comparison.

```{r constraints_dif}
# Gender-stratified separate-group calibration
# gender from bfi$gender is aligned to items_rc row-for-row (both N = 2436)
gender <- bfi$gender[complete.cases(bfi[, 1:25])]   # 1 = male, 2 = female

mod_male   <- TAM::tam.mml(items_rc[gender == 1, c("E1","E2","E3","E4","E5")],
                            verbose = FALSE)
mod_female <- TAM::tam.mml(items_rc[gender == 2, c("E1","E2","E3","E4","E5")],
                            verbose = FALSE)

# Flag items where location differs by more than 0.5 logits between groups
dif_delta  <- mod_male$item[, "xsi.item"] - mod_female$item[, "xsi.item"]
cat("E item DIF (male - female location):\n")
print(round(dif_delta, 3))
flagged_dif <- which(abs(dif_delta) > 0.5)
if (length(flagged_dif)) {
  cat("DIF-flagged items:", paste(names(flagged_dif), collapse = ", "), "\n")
}

# For formal DIF tests: difR::difMH() (binary) or lordif::lordif() (ordinal)
# install.packages(c("difR", "lordif"))
```

---

## 16. Calibration Scores vs. Individual Scores: A Critical Distinction

### The core distinction

`TAM::tam.mml()` uses **marginal maximum likelihood** -- it integrates over the
latent theta distribution to estimate item parameters. The algorithm produces
accurate *item* parameters; it does not produce individual ability estimates
as a byproduct.

Two tools that *look* like individual scores but are not appropriate for individual
reporting:

- **Plausible values (PVs)** are random draws from the posterior
  $\tilde{\theta} \sim p(\theta \mid \mathbf{y}, \hat{\boldsymbol{\xi}})$.
  They are designed to preserve *population* variance for group-level inference
  (PISA, NAEP, TIMSS). Reporting a single PV to a respondent as "your score"
  misrepresents the number -- it could be anywhere in a +-2 SE band around
  their true theta.

- **MI-imputed responses under MAR** similarly yield unbiased *sample* parameter
  estimates; they are not recoveries of what an individual would have answered.
  An ability estimate derived solely from imputed responses should not be
  presented as that person's true standing.

> **Rule of thumb**: for *group-level* statistics (means, variances, regression
> coefficients) use plausible values and pool across the PV set following
> Rubin's rules. For *individual* score reporting, use EAP or WLE.

### What inrep uses for individual reporting

| Estimator | Method | Properties |
|-----------|--------|------------|
| **EAP** | $\hat{\theta} = E[\theta \mid \mathbf{y}]$ | Shrunk toward prior; lower RMSE; best default |
| **WLE** | Warm's Weighted Likelihood Estimate | Approximately unbiased; better for extreme scores |

Both give a point estimate plus a standard error. The SE is what `min_SEM` in
`create_study_config()` monitors. Always report the SE alongside the estimate --
never more than 2 decimal places (false precision for theta scores).

```{r individual_scores}
# inrep's deployment-facing scorer works directly with the calibrated bank
rv_example <- list(
  responses = c(4, 3, 5, 4, 2),
  administered = 1:5,
  current_ability = 0
)
ability_est <- estimate_ability(rv_example, extra_bank, config_extra)
str(ability_est)
# Returns a list with theta and se

# For calibration diagnostics, you can also extract person estimates from the
# TAM model object itself:
person_pars <- TAM::tam.wle(mod_E_anchored)
head(round(person_pars[, c("theta", "error")], 3))
# theta = WLE estimate    error = standard error of measurement
```

```{r plausible_values_group}
# Group-level inference only: generate plausible values, then pool
# mod_E_anchored was fitted on items_rc (N = 2436); pvs and gender are row-aligned
pvs <- TAM::tam.pv(mod_E_anchored, nplausible = 5)$pv

# Pool across PV columns (Rubin's rules); never pick a single PV
for (pv_col in grep("^PV", names(pvs), value = TRUE)) {
  cat(pv_col, "- group means:",
      round(tapply(pvs[[pv_col]], gender, mean, na.rm = TRUE), 3), "\n")
}
# Average the 5 group means and pool their SEs via Rubin's rules for inference.
```

### Practical workflow summary

| Phase | Task | Method |
|-------|------|--------|
| Calibration | Estimate item parameters | `tam.mml()` on calibration sample |
| Deployment | Report individual scores | EAP or WLE from `estimate_ability()` |
| Post-hoc group analysis | Compare group means / regression | Plausible values + Rubin's rules |
