---
title: "Advanced Examples - Cognitive Assessment and Features"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Examples - Cognitive Assessment and Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE  # Set to FALSE for CRAN to avoid Shiny examples
)
```

```{r setup}
library(inrep)
library(TAM)
```

# Advanced Cognitive Assessment Examples

This vignette demonstrates advanced features of the `inrep` package using comprehensive cognitive assessment scenarios.

## Example 1: Basic Personality Assessment (BFI)

```{r basic-bfi}
# Load the built-in Big Five personality items
data(bfi_items)

# Create a simple configuration
basic_config <- create_study_config(
    name = "Personality Assessment",
    model = "GRM",
    max_items = 15,
    min_items = 5,
    criteria = "MI"
)

# Launch the study (requires Shiny environment)
# launch_study(config = basic_config, item_bank = bfi_items)
```

## Example 2: Advanced Cognitive Assessment with 2PL Model

```{r advanced-cognitive}
# Load the cognitive assessment items
data(cognitive_items)

# Create advanced configuration with accessibility features
advanced_config <- create_study_config(
    name = "Cognitive Ability Assessment",
    model = "2PL", 
    estimation_method = "TAM",
    max_items = 20,
    min_items = 10,
    min_SEM = 0.25,
    criteria = "MI",  # Maximum Information selection
    theta_prior = c(0, 1),
    demographics = c("Age", "Gender", "Education", "Native_Language"),
    input_types = list(
        Age = "numeric",
        Gender = "select", 
        Education = "select",
        Native_Language = "text"
    ),
    theme = "Professional",
    session_save = TRUE,
    parallel_computation = TRUE,
    cache_enabled = TRUE,
    accessibility_enhanced = TRUE
)

# Launch with accessibility features
# launch_study(
#     config = advanced_config,
#     item_bank = cognitive_items,
#     accessibility = TRUE
# )
```

## Example 3: Multi-Domain Assessment

```{r multi-domain}
# Create configuration for multiple cognitive domains
multi_domain_config <- create_study_config(
    name = "Comprehensive Cognitive Battery",
    model = "2PL",
    max_items = 30,
    min_items = 15,
    criteria = "MI",
    demographics = c("Age", "Education", "Testing_Experience"),
    input_types = list(
        Age = "numeric",
        Education = "select",
        Testing_Experience = "select"
    ),
    theme = "Academic",
    session_save = TRUE
)

# Show available cognitive domains
print("Available cognitive domains:")
print(unique(cognitive_items$domain))

# Launch multi-domain assessment
# launch_study(
#     config = multi_domain_config,
#     item_bank = cognitive_items
# )
```

## Example 4: Research Study with Cloud Storage

```{r cloud-study}
# Configuration for research study with cloud storage
research_config <- create_study_config(
    name = "Cognitive Research Study 2025",
    model = "GRM",
    max_items = 25,
    min_items = 12,
    criteria = "MI",
    demographics = c("Age", "Gender", "Education", "Handedness"),
    input_types = list(
        Age = "numeric",
        Gender = "select",
        Education = "select",
        Handedness = "select"
    ),
    theme = "Research",
    session_save = TRUE,
    study_key = "COGNITIVE_2025"
)

# WebDAV configuration for cloud storage
webdav_url <- "https://your-cloud-storage.com/webdav"
study_password <- "secure_password"

# Launch with cloud storage (requires valid WebDAV setup)
# launch_study(
#     config = research_config,
#     item_bank = cognitive_items,
#     save_format = "json",
#     webdav_url = webdav_url,
#     webdav_password = study_password
# )
```

## Key Features Demonstrated

### 1. **Advanced Model Configuration**
- Support for 1PL, 2PL, 3PL, and GRM models
- TAM integration for reliable psychometric computation
- Flexible stopping criteria (minimum SEM, maximum items)

### 2. **Comprehensive Demographics**
- Multiple input types (numeric, select, text)
- Customizable demographic questions
- Automatic validation and storage

### 3. **Professional Themes**
- Built-in themes: Light, Dark, Professional, Academic, Research
- Accessibility compliance features
- Responsive design for multiple devices

### 4. **Research Features**
- Session management with pause/resume capability
- Cloud storage integration via WebDAV
- Comprehensive data export formats
- Real-time progress monitoring

### 5. **Cognitive Assessment Domains**
- **Verbal Reasoning**: Language comprehension and analogies
- **Numerical Reasoning**: Mathematical problem solving
- **Spatial Reasoning**: Visual-spatial pattern recognition
- **Working Memory**: Short-term memory and manipulation
- **Processing Speed**: Rapid cognitive processing tasks

## Dataset Information

### BFI Items (Big Five Inventory)
- **Size**: 30 items across 5 personality dimensions
- **Model**: Graded Response Model (GRM) with 5 response categories
- **Domains**: Openness, Conscientiousness, Extraversion, Agreeableness, Neuroticism

### Cognitive Items
- **Size**: 50 items across 5 cognitive domains  
- **Model**: 2PL with calibrated difficulty and discrimination parameters
- **Domains**: Verbal, Numerical, Spatial, Working Memory, Processing Speed

## Technical Notes

All psychometric computations are performed using the **TAM** package, ensuring:
- Reliable parameter estimation
- Standard error calculations
- Information function computations
- Adaptive item selection algorithms

For production deployments, ensure proper citation of both `inrep` and `TAM` packages.

## Next Steps

1. **Customize item banks**: Create domain-specific items for your research
2. **Configure themes**: Match your institutional branding  
3. **Set up cloud storage**: Enable data collection across sessions
4. **Plan deployment**: Choose appropriate hosting platform

For technical support and advanced configuration options, visit the [inrep GitHub repository](https://github.com/selvastics/inrep).
