---
title: "Comprehensive Guide to inrep: Adaptive Testing with Multiple Item Banks"
author: "inrep Package Team"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: united
vignette: >
  %\VignetteIndexEntry{Comprehensive Guide to inrep: Adaptive Testing with Multiple Item Banks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  warning = FALSE,
  message = FALSE
)
```

# Introduction

The `inrep` package provides a comprehensive framework for adaptive testing using Item Response Theory (IRT) models. This vignette demonstrates how to use the package with multiple item banks including personality assessment (Big Five Inventory) and cognitive assessment (Mathematics).

## Package Features

- **Multiple Item Banks**: BFI personality items and mathematics questions
- **Adaptive Testing**: Dynamic item selection based on ability estimates
- **IRT Integration**: Seamless integration with TAM package for psychometric computations
- **Flexible Configuration**: Support for various testing scenarios
- **Interactive Interface**: Shiny-based web application for test administration

## Installation and Setup

```{r eval=FALSE}
# Install from local source
devtools::install_local("path/to/inrep")

# Load the package
library(inrep)
library(TAM)
library(shiny)
```

```{r echo=FALSE}
# For vignette building, we'll simulate loading
library(TAM)
```

# Dataset Overview

The `inrep` package includes several comprehensive datasets for different testing scenarios.

## BFI Items Dataset

The Big Five Inventory (BFI) dataset contains personality assessment items measuring five major personality dimensions:

```{r eval=FALSE}
# Load BFI items
data(bfi_items)
head(bfi_items)

# Dataset structure
str(bfi_items)

# Content areas
table(bfi_items$Content_Area)
```

**BFI Content Areas:**
- **Extraversion**: Sociability, assertiveness, activity level
- **Agreeableness**: Trust, cooperation, empathy
- **Conscientiousness**: Organization, responsibility, dependability
- **Neuroticism**: Emotional stability, anxiety, mood
- **Openness**: Creativity, intellectual curiosity, aesthetic appreciation

## Mathematics Items Dataset

The mathematics dataset contains entry-level mathematics questions across multiple content areas:

```{r eval=FALSE}
# Load mathematics items
data(math_items)
head(math_items)

# Content areas
table(math_items$Content_Area)
table(math_items$Difficulty_Level)
```

**Mathematics Content Areas:**
- **Arithmetic**: Basic operations, percentages, fractions
- **Algebra**: Linear equations, simplification, substitution
- **Geometry**: Area, perimeter, volume calculations
- **Statistics**: Mean, median, probability, range

## Combined Dataset

For comprehensive assessment scenarios, items can be combined across domains:

```{r eval=FALSE}
# Create combined assessment
combined_items <- rbind(
  transform(bfi_items, Domain = "Personality"),
  transform(math_items, Domain = "Cognitive")
)

# Overview
table(combined_items$Domain, combined_items$Content_Area)
```

# Basic Usage Examples

## Example 1: Simple BFI Assessment

```{r eval=FALSE}
# Create study configuration for BFI assessment
config <- create_study_config(
  study_name = "BFI Personality Assessment",
  item_bank = "bfi_items",
  estimation_method = "TAM",
  max_items = 20,
  min_items = 10,
  stopping_criterion = "se_threshold",
  se_threshold = 0.5,
  demographics = list(
    age = list(type = "numeric", required = TRUE),
    gender = list(type = "select", options = c("Male", "Female", "Other")),
    education = list(type = "select", options = c("High School", "Bachelor", "Master", "PhD"))
  )
)

# Launch the assessment
launch_study(config)
```

## Example 2: Mathematics Adaptive Test

```{r eval=FALSE}
# Create mathematics adaptive test
math_config <- create_study_config(
  study_name = "Mathematics Adaptive Test",
  item_bank = "math_items",
  estimation_method = "TAM",
  max_items = 15,
  min_items = 5,
  stopping_criterion = "se_threshold",
  se_threshold = 0.4,
  adaptive_selection = TRUE,
  selection_method = "maximum_info",
  demographics = list(
    grade_level = list(type = "select", options = c("6th", "7th", "8th", "9th", "10th")),
    math_background = list(type = "select", options = c("Basic", "Intermediate", "Advanced"))
  )
)

# Launch the test
launch_study(math_config)
```

## Example 3: Combined Assessment

```{r eval=FALSE}
# Create comprehensive assessment combining both domains
comprehensive_config <- create_study_config(
  study_name = "Comprehensive Assessment: Personality & Cognitive",
  item_bank = "combined_items",
  estimation_method = "TAM",
  max_items = 30,
  min_items = 15,
  stopping_criterion = "max_items",
  adaptive_selection = TRUE,
  selection_method = "balanced_info",
  demographics = list(
    age = list(type = "numeric", required = TRUE),
    education = list(type = "select", options = c("High School", "Bachelor", "Master", "PhD")),
    research_participation = list(type = "checkbox", required = TRUE)
  )
)

# Launch comprehensive assessment
launch_study(comprehensive_config)
```

# Advanced Configuration Options

## Adaptive Testing Parameters

The package supports sophisticated adaptive testing configurations:

```{r eval=FALSE}
# Advanced adaptive configuration
adaptive_config <- create_study_config(
  study_name = "Advanced Adaptive Test",
  item_bank = "math_items",
  estimation_method = "TAM",
  
  # Adaptive parameters
  adaptive_selection = TRUE,
  selection_method = "maximum_info",
  exposure_control = TRUE,
  max_exposure_rate = 0.3,
  content_balancing = TRUE,
  
  # Stopping criteria
  stopping_criterion = "multiple",
  max_items = 20,
  min_items = 8,
  se_threshold = 0.35,
  theta_change_threshold = 0.1,
  
  # Initial parameters
  initial_theta = 0,
  initial_se = 1,
  
  # Demographics
  demographics = list(
    participant_id = list(type = "text", required = TRUE),
    test_date = list(type = "date", required = TRUE),
    testing_conditions = list(type = "select", options = c("Laboratory", "Online", "Classroom"))
  )
)
```

## Custom Item Selection

```{r eval=FALSE}
# Custom item selection function
custom_selection <- function(theta, item_bank, administered_items) {
  # Implement custom logic for item selection
  # This could include:
  # - Content area rotation
  # - Difficulty progression
  # - Exposure control
  # - Domain-specific rules
  
  available_items <- setdiff(1:nrow(item_bank), administered_items)
  
  # Example: Select item with maximum information at current theta
  info_values <- sapply(available_items, function(i) {
    item_info(theta, item_bank$a[i], item_bank$b[i])
  })
  
  return(available_items[which.max(info_values)])
}

# Use custom selection in configuration
config_custom <- create_study_config(
  study_name = "Custom Selection Test",
  item_bank = "math_items",
  selection_function = custom_selection,
  # ... other parameters
)
```

# Analysis and Reporting

## Ability Estimation

```{r eval=FALSE}
# Estimate abilities from response data
responses <- c(1, 0, 1, 1, 0, 1, 0, 1)  # Example responses
item_indices <- c(1, 3, 5, 7, 9, 11, 13, 15)  # Administered items

ability_est <- estimate_ability(
  responses = responses,
  item_indices = item_indices,
  item_bank = "math_items",
  method = "ML"
)

print(ability_est)
```

## Test Information

```{r eval=FALSE}
# Calculate test information function
theta_range <- seq(-3, 3, by = 0.1)
test_info <- calculate_test_info(
  theta_range = theta_range,
  item_bank = "math_items",
  administered_items = item_indices
)

# Plot test information
plot(theta_range, test_info, type = "l", 
     xlab = "Ability (θ)", ylab = "Test Information",
     main = "Test Information Function")
```

## Response Pattern Analysis

```{r eval=FALSE}
# Analyze response patterns
response_analysis <- analyze_responses(
  responses = responses,
  item_indices = item_indices,
  item_bank = "math_items"
)

# View analysis results
print(response_analysis)
```

# Practical Applications

## Educational Assessment

```{r eval=FALSE}
# Create educational assessment battery
education_config <- create_study_config(
  study_name = "Educational Assessment Battery",
  item_bank = "math_items",
  estimation_method = "TAM",
  max_items = 25,
  min_items = 12,
  stopping_criterion = "se_threshold",
  se_threshold = 0.3,
  adaptive_selection = TRUE,
  content_balancing = TRUE,
  demographics = list(
    student_id = list(type = "text", required = TRUE),
    grade = list(type = "select", options = c("6", "7", "8", "9", "10", "11", "12")),
    school = list(type = "text", required = TRUE),
    teacher = list(type = "text", required = FALSE)
  )
)

# Launch educational assessment
launch_study(education_config)
```

## Psychological Research

```{r eval=FALSE}
# Create research study configuration
research_config <- create_study_config(
  study_name = "Personality Research Study",
  item_bank = "bfi_items",
  estimation_method = "TAM",
  max_items = 44,  # All BFI items
  min_items = 20,
  stopping_criterion = "max_items",
  adaptive_selection = FALSE,  # Fixed-length administration
  randomize_items = TRUE,
  demographics = list(
    participant_id = list(type = "text", required = TRUE),
    age = list(type = "numeric", required = TRUE, min = 18, max = 80),
    gender = list(type = "select", options = c("Male", "Female", "Non-binary", "Prefer not to say")),
    education = list(type = "select", options = c("High School", "Some College", "Bachelor", "Master", "PhD")),
    consent = list(type = "checkbox", required = TRUE, label = "I consent to participate in this research")
  )
)

# Launch research study
launch_study(research_config)
```

## Clinical Assessment

```{r eval=FALSE}
# Create clinical assessment configuration
clinical_config <- create_study_config(
  study_name = "Clinical Psychological Assessment",
  item_bank = "bfi_items",
  estimation_method = "TAM",
  max_items = 30,
  min_items = 15,
  stopping_criterion = "se_threshold",
  se_threshold = 0.4,
  adaptive_selection = TRUE,
  selection_method = "maximum_info",
  demographics = list(
    client_id = list(type = "text", required = TRUE),
    age = list(type = "numeric", required = TRUE),
    referral_source = list(type = "select", options = c("Self", "Physician", "Therapist", "Other")),
    assessment_date = list(type = "date", required = TRUE)
  )
)

# Launch clinical assessment
launch_study(clinical_config)
```

# Customization and Extensions

## Custom Themes

```{r eval=FALSE}
# Create custom visual theme
custom_theme <- create_custom_theme(
  primary_color = "#2E86AB",
  secondary_color = "#A23B72",
  background_color = "#F18F01",
  text_color = "#333333",
  font_family = "Arial",
  font_size = "14px"
)

# Apply theme to study
themed_config <- create_study_config(
  study_name = "Themed Assessment",
  item_bank = "math_items",
  theme = custom_theme,
  # ... other parameters
)
```

## Custom Validation

```{r eval=FALSE}
# Create custom validation rules
custom_validation <- function(item_bank) {
  # Custom validation logic
  errors <- character(0)
  
  # Check for required columns
  required_cols <- c("Question", "Option1", "Option2", "Option3", "Option4", "Answer", "a", "b")
  missing_cols <- setdiff(required_cols, names(item_bank))
  if (length(missing_cols) > 0) {
    errors <- c(errors, paste("Missing columns:", paste(missing_cols, collapse = ", ")))
  }
  
  # Check IRT parameters
  if (any(item_bank$a <= 0)) {
    errors <- c(errors, "Discrimination parameters must be positive")
  }
  
  # Check answer validity
  valid_answers <- c("Option1", "Option2", "Option3", "Option4")
  if (!all(item_bank$Answer %in% valid_answers)) {
    errors <- c(errors, "Invalid answer options found")
  }
  
  return(errors)
}

# Use custom validation
validated_config <- create_study_config(
  study_name = "Validated Assessment",
  item_bank = "math_items",
  validation_function = custom_validation,
  # ... other parameters
)
```

# Best Practices

## Test Administration

1. **Pilot Testing**: Always pilot test your configuration with a small sample
2. **Item Exposure**: Monitor item exposure rates to ensure test security
3. **Content Balance**: Ensure adequate representation across content areas
4. **Stopping Rules**: Use appropriate stopping criteria for your context

## Data Quality

1. **Response Time**: Monitor response times for quality control
2. **Response Patterns**: Check for aberrant response patterns
3. **Missing Data**: Handle missing responses appropriately
4. **Validation**: Validate item bank data before use

## Ethical Considerations

1. **Informed Consent**: Ensure participants understand the assessment
2. **Data Privacy**: Protect participant data and responses
3. **Feedback**: Provide appropriate feedback to participants
4. **Accessibility**: Ensure tests are accessible to all participants

# Troubleshooting

## Common Issues

1. **Item Bank Errors**: Check data format and column names
2. **TAM Integration**: Ensure TAM package is installed and loaded
3. **Shiny Interface**: Check for JavaScript errors in browser console
4. **Memory Issues**: Monitor memory usage with large item banks

## Error Messages

- `"Item bank validation failed"`: Check item bank format and content
- `"TAM estimation failed"`: Check response data and item parameters
- `"Configuration error"`: Review study configuration parameters
- `"Shiny app launch failed"`: Check Shiny dependencies and port availability

# Conclusion

The `inrep` package provides a comprehensive framework for adaptive testing with multiple item banks. Whether you're conducting educational assessments, psychological research, or clinical evaluations, the package offers the flexibility and power needed for sophisticated testing scenarios.

For more information, visit the package documentation or contact the development team.

## References

1. Hambleton, R. K., Swaminathan, H., & Rogers, H. J. (1991). *Fundamentals of item response theory*. Sage Publications.

2. van der Linden, W. J., & Glas, C. A. (Eds.). (2010). *Elements of adaptive testing*. Springer.

3. Robitzsch, A., Kiefer, T., & Wu, M. (2020). *TAM: Test Analysis Modules*. R package version 3.5-19.

4. John, O. P., & Srivastava, S. (1999). The Big Five trait taxonomy: History, measurement, and theoretical perspectives. *Handbook of personality: Theory and research*, 2(1999), 102-138.
