---
title: "Customizing Appearance and Themes"
author: "Clievins Selva"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customizing Appearance and Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

This vignette explains how to customize the appearance of your adaptive tests using **inrep**'s theming system and website scraping capabilities. You'll learn how to create professional, branded assessments that match your institution's visual identity.

## Built-in Themes

**inrep** includes several professionally designed themes:

```{r}
library(inrep)

# View available themes
get_builtin_themes()
```

### Professional Theme

Clean, corporate appearance suitable for business assessments:

```{r}
config_professional <- create_study_config(
  name = "Leadership Assessment",
  theme = "professional",
  colors = list(
    primary = "#2E5984",
    secondary = "#8EADC7",
    background = "#FFFFFF",
    text = "#333333"
  )
)
```

### Academic Theme

Traditional academic styling for research studies:

```{r}
config_academic <- create_study_config(
  name = "Cognitive Research Study",
  theme = "academic",
  colors = list(
    primary = "#8B0000",
    secondary = "#CD853F",
    background = "#FFFEF7"
  )
)
```

### Modern Theme

Contemporary design with clean lines:

```{r}
config_modern <- create_study_config(
  name = "Innovation Assessment",
  theme = "modern",
  colors = list(
    primary = "#4A90E2",
    secondary = "#7ED321",
    accent = "#F5A623"
  )
)
```

## Website Theme Scraping

**inrep** can automatically extract colors, fonts, and styling from institutional websites to create matching themes.

### Basic Website Scraping

```{r}
# Scrape university website for theme
result <- scrape_website_ui("https://www.uni-hildesheim.de/")

# View extracted themes
for (i in seq_along(result$themes)) {
  theme <- result$themes[[i]]
  cat(sprintf("Theme %d: %s (Primary: %s)\\n", 
              i, theme$name, theme$primary_color))
}
```

### Using Scraped Themes

```{r}
# Use the first extracted theme
config_scraped <- create_study_config(
  name = "University Research Study",
  theme_config = result$themes[[1]]
)

# Launch with multiple theme options
launch_study(config_scraped, bfi_items, theme_options = result$themes)
```

### Advanced Scraping Options

```{r}
# Scrape with specific targeting
result_advanced <- scrape_website_ui(
  url = "https://www.university.edu/",
  extract_options = list(
    target_elements = c("header", "navigation", "main"),
    include_fonts = TRUE,
    include_logos = TRUE,
    color_palette_size = 5
  )
)
```

## Custom Theme Creation

### Manual Theme Configuration

Create completely custom themes:

```{r}
custom_theme <- list(
  name = "Corporate Blue",
  primary_color = "#0066CC",
  secondary_color = "#99CCFF", 
  background_color = "#F8F9FA",
  text_color = "#212529",
  accent_color = "#FF6B35",
  
  fonts = list(
    heading = "Roboto Slab",
    body = "Open Sans",
    monospace = "Source Code Pro"
  ),
  
  styling = list(
    border_radius = "8px",
    box_shadow = "0 2px 4px rgba(0,0,0,0.1)",
    button_style = "solid",
    input_style = "outlined"
  )
)

config_custom <- create_study_config(
  name = "Custom Branded Assessment",
  theme_config = custom_theme
)
```

### Institution-Specific Branding

```{r}
# University branding example
uni_theme <- list(
  name = "University of Excellence",
  primary_color = "#8B0000",        # University red
  secondary_color = "#FFD700",      # University gold
  background_color = "#FFFFFF",
  
  logo = list(
    url = "https://university.edu/logo.png",
    position = "header-left",
    max_height = "60px"
  ),
  
  fonts = list(
    heading = "Georgia",            # Serif for headings
    body = "Arial"                  # Sans-serif for body
  ),
  
  footer = list(
    text = "© 2025 University of Excellence",
    links = list(
      "Privacy Policy" = "https://university.edu/privacy",
      "Contact" = "mailto:research@university.edu"
    )
  )
)
```

## Advanced Customization

### CSS Customization

For complete control, provide custom CSS:

```{r}
config_css <- create_study_config(
  name = "Fully Custom Assessment",
  custom_css = "
    .assessment-container {
      background: linear-gradient(135deg, #667eea 0%, #764ba2 100%);
      min-height: 100vh;
    }
    
    .question-card {
      backdrop-filter: blur(10px);
      background: rgba(255, 255, 255, 0.9);
      border-radius: 15px;
      box-shadow: 0 8px 32px rgba(31, 38, 135, 0.37);
    }
    
    .btn-primary {
      background: linear-gradient(45deg, #667eea, #764ba2);
      border: none;
      transition: transform 0.2s;
    }
    
    .btn-primary:hover {
      transform: translateY(-2px);
    }
  "
)
```

### Dynamic Theme Switching

Allow participants to choose themes:

```{r}
# Multiple theme options
theme_light <- list(name = "Light", primary_color = "#007bff")
theme_dark <- list(name = "Dark", primary_color = "#343a40", 
                   background_color = "#212529", text_color = "#ffffff")
theme_high_contrast <- list(name = "High Contrast", 
                           primary_color = "#000000", 
                           background_color = "#ffffff")

config_multi <- create_study_config(
  name = "Accessible Assessment",
  theme_options = list(theme_light, theme_dark, theme_high_contrast),
  allow_theme_switching = TRUE
)
```

## Accessibility Considerations

### High Contrast Themes

```{r}
accessible_theme <- list(
  name = "High Contrast",
  primary_color = "#000000",
  secondary_color = "#FFFFFF", 
  background_color = "#FFFFFF",
  text_color = "#000000",
  
  accessibility = list(
    high_contrast = TRUE,
    large_fonts = TRUE,
    focus_indicators = "enhanced",
    screen_reader_optimized = TRUE
  )
)
```

### Font Size Options

```{r}
config_accessible <- create_study_config(
  name = "Accessible Assessment",
  accessibility_features = list(
    font_size_options = c("small", "medium", "large", "extra-large"),
    default_font_size = "medium",
    dyslexia_friendly_font = TRUE
  )
)
```

## Responsive Design

Themes automatically adapt to different screen sizes:

```{r}
responsive_config <- create_study_config(
  name = "Mobile-Friendly Assessment",
  responsive_design = list(
    mobile_optimized = TRUE,
    tablet_layout = "adaptive",
    desktop_layout = "full-width"
  ),
  
  mobile_settings = list(
    swipe_navigation = TRUE,
    touch_friendly_buttons = TRUE,
    auto_zoom_prevention = TRUE
  )
)
```

## Theme Testing and Preview

### Preview Themes

```{r}
# Launch theme editor for testing
launch_theme_editor(
  themes = list(theme_light, theme_dark, custom_theme),
  sample_content = "assessment_preview"
)
```

### A/B Testing Themes

```{r}
# Test multiple themes with participants
config_ab_test <- create_study_config(
  name = "Theme Comparison Study",
  ab_test_themes = list(
    control = theme_light,
    treatment = theme_dark
  ),
  ab_test_ratio = 0.5  # 50/50 split
)
```

## Integration Examples

### Corporate Assessment

```{r}
# Fortune 500 company assessment
corporate_config <- create_study_config(
  name = "Leadership Development Assessment",
  
  theme_config = list(
    name = "Corporate Excellence",
    primary_color = "#003366",      # Navy blue
    secondary_color = "#0066CC",    # Light blue
    accent_color = "#FF6600",       # Orange accent
    
    logo = list(
      url = "https://company.com/logo.svg",
      position = "header-center"
    ),
    
    branding = list(
      company_name = "ExampleCorp",
      tagline = "Excellence in Leadership",
      footer_text = "Confidential Assessment"
    )
  ),
  
  styling = list(
    professional_layout = TRUE,
    progress_indicator = "bar",
    question_numbering = "hidden"
  )
)
```

### Academic Research

```{r}
# University research study
research_config <- create_study_config(
  name = "Cognitive Psychology Research",
  
  theme_config = list(
    name = "Academic Research",
    primary_color = "#8B0000",
    background_color = "#FFFEF7",
    
    header = list(
      title = "Department of Psychology",
      subtitle = "Research Study Participation",
      institutional_logo = TRUE
    )
  ),
  
  research_features = list(
    participant_id_visible = FALSE,
    progress_hidden = TRUE,
    neutral_styling = TRUE
  )
)
```

### Clinical Assessment

```{r}
# Healthcare/clinical setting
clinical_config <- create_study_config(
  name = "Patient Assessment Tool",
  
  theme_config = list(
    name = "Healthcare Professional",
    primary_color = "#0F4C81",      # Medical blue
    secondary_color = "#E8F4FD",    # Light blue
    
    clinical_styling = list(
      clean_interface = TRUE,
      minimal_distractions = TRUE,
      professional_fonts = TRUE
    )
  ),
  
  privacy_features = list(
    no_external_resources = TRUE,
    secure_styling = TRUE,
    hipaa_compliant = TRUE
  )
)
```

## Best Practices

### Theme Selection Guidelines

1. **Professional Assessments**: Use corporate colors and clean typography
2. **Research Studies**: Prefer neutral, academic themes
3. **Educational Tests**: Choose engaging but not distracting themes
4. **Clinical Tools**: Prioritize clarity and accessibility

### Performance Considerations

```{r}
# Optimize for performance
optimized_config <- create_study_config(
  name = "High-Performance Assessment",
  
  performance_optimization = list(
    minimize_css = TRUE,
    lazy_load_images = TRUE,
    cache_fonts = TRUE,
    compress_assets = TRUE
  )
)
```

### Cross-Browser Compatibility

Themes are tested across major browsers and automatically include fallbacks for older browsers.

## Conclusion

**inrep**'s theming system provides powerful tools for creating professional, branded assessments. Whether using built-in themes, scraping institutional websites, or creating custom designs, you can ensure your assessments match your organization's visual identity while maintaining accessibility and usability.

For more advanced customization options, see the "Research Workflows" vignette.
