---
title: "Getting Started with inrep: Your First Adaptive Test"
author: "inrep Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with inrep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
library(inrep)
```

# Welcome to inrep!

The **inrep** package is a comprehensive framework for creating and administering adaptive assessments using Item Response Theory (IRT) models. Whether you're an educator, researcher, or assessment professional, inrep provides the tools you need to create sophisticated adaptive tests with minimal effort.

## What Makes inrep Special?

- **Multiple IRT Models**: Support for 1PL, 2PL, 3PL, and Graded Response Model (GRM)
- **Dual Estimation Engines**: Choose between TAM and MIRT for ability estimation
- **Adaptive & Fixed Testing**: Flexible testing modes for different scenarios
- **Beautiful UI**: Customizable themes and multilingual support
- **Research Ready**: Comprehensive logging and data export capabilities
- **CRAN Quality**: Production-ready code with extensive documentation

## Installation

```{r eval=FALSE}

# Install development version from GitHub
devtools::install_github("your-repo/inrep")
```

## Quick Start: Your First Adaptive Test

Let's create a simple adaptive test in just a few lines of code:

```{r basic_example, eval=FALSE}
library(inrep)

# Use the built-in personality item bank (automatically loaded via LazyData)
# bfi_items is available when the package is loaded

# Create a basic study configuration
config <- create_study_config(
  name = "My First Adaptive Test",
  model = "2PL",                    # Two-parameter logistic model
  estimation_method = "TAM",        # Use TAM for estimation
  max_items = 10,                   # Maximum 10 items
  min_items = 5,                    # Minimum 5 items
  criteria = "MI"                   # Maximum Information criterion
)

# Launch the adaptive test
launch_study(config, bfi_items)
```

That's it! This code creates a fully functional adaptive test interface.

## Understanding the Components

### 1. Item Banks

Item banks contain the questions and their psychometric properties. Let's examine the built-in item bank:

```{r examine_items}
# Examine the structure (data already loaded above)
head(bfi_items, 3)
cat("Number of items:", nrow(bfi_items), "\n")
cat("Available columns:", names(bfi_items), "\n")
```

**Key columns explained:**
- `Question`: The item text displayed to users
- `a`: Discrimination parameter (how well the item differentiates)
- `b`: Difficulty/location parameter 
- `Option1-Option4`: Response options for the item
- Additional metadata like `Trait`, `Domain`, etc.

### 2. Study Configuration

The configuration object controls all aspects of your adaptive test:

```{r detailed_config}
# Create a more detailed configuration
advanced_config <- create_study_config(
  name = "Advanced Personality Assessment",
  model = "2PL",
  estimation_method = "TAM",        # Use TAM (MIRT will be added later)
  max_items = 20,
  min_items = 8,
  min_SEM = 0.3,                    # Stop when standard error < 0.3
  criteria = "MI",                  # Maximum Information
  adaptive = TRUE,                  # Enable adaptive selection
  theta_prior = c(0, 1),           # Prior: mean=0, sd=1
  demographics = c("Age", "Gender") # Collect demographics
)

# View the configuration
str(advanced_config)
```

### 3. IRT Models Available

inrep supports four major IRT models:

```{r models_demo, eval=FALSE}
# 1-Parameter Logistic (Rasch) Model - Equal discrimination
config_1pl <- create_study_config(
  name = "Rasch Model Test",
  model = "1PL",
  estimation_method = "TAM"
)

# 2-Parameter Logistic Model - Variable discrimination
config_2pl <- create_study_config(
  name = "2PL Model Test", 
  model = "2PL",
  estimation_method = "TAM"
)

# 3-Parameter Logistic Model - Includes guessing
config_3pl <- create_study_config(
  name = "3PL Model Test",
  model = "3PL", 
  estimation_method = "TAM"
)

# Graded Response Model - For polytomous items
config_grm <- create_study_config(
  name = "GRM Likert Scale Test",
  model = "GRM",
  estimation_method = "TAM"
)
```

## Estimation Methods: TAM vs MIRT

inrep offers two powerful estimation engines:

### TAM (Test Analysis Modules)
- **Best for**: Quick, reliable estimates
- **Advantages**: Fast, well-tested, stable
- **Recommended**: Educational testing, large-scale assessments

### MIRT (Multidimensional IRT)
- **Best for**: Research applications, complex models
- **Advantages**: More accurate, better diagnostics, polytomous support
- **Recommended**: Psychological research, detailed analysis

```{r estimation_comparison, eval=FALSE}
# Compare estimation methods
set.seed(123)

# Simulate some responses
responses <- c(1, 0, 1, 1, 0, 1, 0, 1)
administered <- 1:8

# TAM estimation
tam_result <- estimate_ability(
  rv = list(responses = responses, administered = administered),
  item_bank = bfi_items,
  config = create_study_config(estimation_method = "TAM")
)

# MIRT estimation  
mirt_result <- estimate_ability_mirt(
  responses = responses,
  administered = administered,
  item_bank = bfi_items,
  model = "2PL",
  method = "EAP"
)

cat("TAM estimate:", sprintf("θ = %.3f (SE = %.3f)", tam_result$theta, tam_result$se), "\n")
cat("MIRT estimate:", sprintf("θ = %.3f (SE = %.3f)", mirt_result$theta, mirt_result$se), "\n")
```

## Customizing Your Test Experience

### Demographics Collection

```{r demographics_demo, eval=FALSE}
config_with_demographics <- create_study_config(
  name = "Research Study",
  demographics = c("Age", "Gender", "Education", "Experience"),
  input_types = list(
    Age = "numeric",
    Gender = "select", 
    Education = "select",
    Experience = "text"
  )
)
```

### Custom Stopping Rules

```{r stopping_rules, eval=FALSE}
# Create a custom stopping rule
custom_stopping <- function(rv, item_bank, config) {
  # Stop if we have 10+ items AND standard error < 0.4
  if (length(rv$administered) >= 10 && rv$current_se < 0.4) {
    return(TRUE)
  }
  # Or if we've reached maximum items
  if (length(rv$administered) >= config$max_items) {
    return(TRUE)
  }
  return(FALSE)
}

config_custom_stop <- create_study_config(
  name = "Custom Stopping Rule Test",
  stopping_rule = custom_stopping,
  max_items = 25
)
```

## Testing Your Setup

Before deploying your adaptive test, validate your configuration:

```{r validation}
# Validate item bank
validation_result <- validate_item_bank(bfi_items, "GRM")
cat("Item bank validation:", ifelse(validation_result, "PASSED", "FAILED"), "\n")

# Check configuration
config_test <- create_study_config(
  name = "Validation Test",
  model = "2PL",
  max_items = 10
)

cat("Configuration created successfully!\n")
cat("Study name:", config_test$name, "\n") 
cat("Model:", config_test$model, "\n")
cat("Max items:", config_test$max_items, "\n")
```

## Next Steps

Now that you understand the basics:

1. **Try the examples**: Run the code snippets above
2. **Explore themes**: Check out `vignette("customizing-appearance")`
3. **Learn advanced features**: See `vignette("advanced-adaptive-testing")`
4. **Research applications**: Read `vignette("research-workflows")`
5. **Deployment**: Learn about `vignette("production-deployment")`

## Common Patterns

Here are some typical use cases to get you started:

### Educational Assessment
```{r educational_pattern, eval=FALSE}
education_config <- create_study_config(
  name = "Math Proficiency Test",
  model = "1PL",                    # Rasch model for education
  estimation_method = "TAM",        # Fast, reliable
  max_items = 15,
  min_items = 10,
  min_SEM = 0.32,                   # Educational standard
  criteria = "MI"
)
```

### Psychological Research
```{r psychology_pattern, eval=FALSE}
research_config <- create_study_config(
  name = "Personality Research Study",
  model = "2PL",                    # Better fit for traits
  estimation_method = "TAM",        # Reliable and fast
  max_items = 30,
  min_items = 15,
  demographics = c("Age", "Gender", "Country"),
  session_save = TRUE               # Save for later analysis
)
```

### Clinical Assessment
```{r clinical_pattern, eval=FALSE}
clinical_config <- create_study_config(
  name = "Depression Screening",
  model = "GRM",                    # Polytomous responses
  estimation_method = "TAM",        # Reliable for clinical use
  max_items = 12,
  min_items = 8,
  min_SEM = 0.25,                   # High precision needed
  criteria = "MI"
)
```

## Getting Help

- **Documentation**: `help(package = "inrep")`
- **Function help**: `?launch_study`, `?create_study_config`
- **Vignettes**: `vignette(package = "inrep")`
- **Issues**: Report bugs on GitHub

Welcome to the world of adaptive testing with inrep! Let's get you started with your first adaptive assessment.
