---
title: "Psychological Study on R Package Testing Experience"
author: "inrep Package"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Psychological Study on R Package Testing Experience}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Psychological Study on R Package Testing Experience

This vignette demonstrates how to create a comprehensive psychological study using the `inrep` package, specifically reproducing the "Psychological Study on R Package Testing Experience" with the elegant **Monochrome** theme.

## Overview

The study examines psychological factors affecting R package testing workflows, exploring how tools like `testthat`, `covr`, and `devtools` influence developer confidence, stress, and motivation. This example showcases:

- **Monochrome Theme**: Elegant black and white design optimized for psychological studies
- **Comprehensive Demographics**: Age range and experience collection
- **Likert Scale Questions**: Professional psychological assessment items
- **Timer Integration**: 20-minute time limit with warnings
- **Progress Tracking**: Visual progress indicators
- **Results Download**: Exportable response data
- **Accessibility Features**: WCAG 2.1 AA compliance

## Study Configuration

### 1. Item Bank Creation

First, we create an item bank with the psychological study questions:

```{r create_item_bank}
library(inrep)

# Create the item bank for the psychological study
r_testing_items <- data.frame(
  item_id = 1:6,
  item_text = c(
    "I feel confident in my ability to write effective unit tests using testthat.",
    "Using devtools to manage and test R packages is intuitive and efficient.",
    "Interpreting code coverage reports from covr helps me improve my testing strategy.",
    "Testing R packages reduces my stress when preparing for a package release.",
    "I feel motivated to incorporate regular testing into my R package development process.",
    "The documentation for testthat is clear and supports my testing needs."
  ),
  explanation = c(
    "Confidence in testing skills can enhance code quality and reduce development errors.",
    "Ease of use in testing workflows impacts developer productivity and adoption.",
    "Clear coverage reports enable targeted improvements in test coverage.",
    "Effective testing can alleviate concerns about introducing bugs in production.",
    "Motivation to test regularly correlates with higher code reliability and quality.",
    "Clear documentation is critical for effective use of testing tools."
  ),
  # Likert scale parameters (treating as ordinal for demonstration)
  a = rep(1.5, 6),  # Discrimination parameter
  b1 = rep(-2, 6),  # Threshold 1 (Strongly Disagree -> Disagree)
  b2 = rep(-1, 6),  # Threshold 2 (Disagree -> Neutral)
  b3 = rep(0, 6),   # Threshold 3 (Neutral -> Agree)
  b4 = rep(1, 6),   # Threshold 4 (Agree -> Strongly Agree)
  response_options = I(rep(list(c(
    "Strongly Disagree", "Disagree", "Neutral", "Agree", "Strongly Agree"
  )), 6))
)

# Validate the item bank
validate_item_bank(r_testing_items, model = "GRM")
```

### 2. Study Configuration

Next, we configure the study with comprehensive settings:

```{r create_study_config}
# Create study configuration with Monochrome theme
study_config <- create_study_config(
  name = "Psychological Study on R Package Testing Experience",
  
  # Theme Configuration
  theme = "Monochrome",  # Use the elegant monochrome theme
  
  # Demographics
  demographics = c("age_range", "r_experience"),
  input_types = list(
    age_range = "radio",
    r_experience = "radio"
  ),
  
  # Study Parameters
  study_key = "r_testing_psychology_2025",
  model = "GRM",  # Graded Response Model for Likert scales
  
  # Adaptive Testing (disabled for fixed questionnaire)
  adaptive = FALSE,
  max_items = 6,
  min_items = 6,
  
  # UI Configuration
  response_ui_type = "radio",
  progress_style = "bar",
  
  # Session Management
  session_save = TRUE,
  session_timeout = 20,  # 20 minutes
  
  # Custom validation for required responses
  response_validation_fun = function(response) {
    !is.null(response) && nzchar(as.character(response))
  },
  
  # Instructions and consent
  instructions = list(
    welcome = "Thank you for participating in this psychological study conducted by researchers investigating the experiences of R developers in package testing. This study explores how tools such as testthat, covr, and devtools influence developer confidence, stress, and motivation.",
    
    purpose = "To examine psychological factors affecting R package testing workflows.",
    
    duration = "Approximately 15-20 minutes.",
    
    structure = "The study consists of a consent form, demographic questions, and a main survey with Likert-scale questions. All questions are optional, but complete responses are appreciated.",
    
    confidentiality = "Your responses are anonymous and will be used solely for research purposes. Data will be stored securely and reported in aggregate form.",
    
    consent_text = "This study involves completing a survey about your experiences with R package testing tools. Participation is voluntary, and you may withdraw at any time by closing this window. Your responses will be anonymized and used for academic research only.",
    
    contact = "For questions, contact the research team at research@example.edu."
  ),
  
  # Demographic question configurations
  demographic_configs = list(
    age_range = list(
      question = "What is your age range?",
      options = c("18-24", "25-34", "35-44", "45-54", "55+"),
      required = FALSE
    ),
    r_experience = list(
      question = "How many years have you been using R for package development?",
      options = c("Less than 1 year", "1-3 years", "3-5 years", "5+ years"),
      required = FALSE
    )
  ),
  
  # Custom recommendation function
  recommendation_fun = function(theta, demographics) {
    # Provide personalized feedback based on responses
    if (is.null(theta) || length(theta) == 0) {
      return("Thank you for participating in this psychological study on R package testing.")
    }
    
    avg_score <- mean(theta, na.rm = TRUE)
    
    if (avg_score >= 1) {
      return("Your responses indicate high confidence and positive experiences with R testing tools. Consider sharing your expertise with the community through tutorials or mentoring.")
    } else if (avg_score >= 0) {
      return("Your responses show moderate comfort with R testing tools. Consider exploring advanced testing techniques or joining R testing communities for support.")
    } else {
      return("Your responses suggest room for growth in R testing practices. Consider starting with basic testthat tutorials and gradually building your testing skills.")
    }
  }
)
```

### 3. Enhanced Study Configuration with Timer

Let's add timer functionality and enhanced features:

```{r enhanced_config}
# Enhanced configuration with timer and progress tracking
enhanced_config <- create_study_config(
  name = "Psychological Study on R Package Testing Experience",
  theme = "Monochrome",
  
  # All previous configurations...
  demographics = c("age_range", "r_experience"),
  input_types = list(
    age_range = "radio",
    r_experience = "radio"
  ),
  
  study_key = "r_testing_psychology_2025",
  model = "GRM",
  adaptive = FALSE,
  max_items = 6,
  min_items = 6,
  
  # Timer Configuration
  session_timeout = 20,  # 20 minutes
  show_timer = TRUE,
  timer_warning_threshold = 2,  # Warning when 2 minutes remain
  
  # Progress Configuration
  progress_style = "bar",
  show_progress = TRUE,
  progress_animations = TRUE,
  
  # Results Configuration
  show_results = TRUE,
  allow_results_download = TRUE,
  results_format = "txt",
  
  # Custom UI elements
  custom_ui_pre = function() {
    shiny::div(
      class = "study-introduction",
      shiny::h1("Psychological Study on R Package Testing Experience", 
                class = "study-title"),
      shiny::p("Welcome to this comprehensive psychological study examining developer experiences with R testing tools.",
               class = "instruction-text")
    )
  },
  
  # Custom CSS for Monochrome theme enhancements
  custom_css = "
    .study-introduction {
      text-align: center;
      margin-bottom: 2rem;
      padding: 2rem;
      background: white;
      border-radius: 6px;
      box-shadow: 0 4px 8px rgba(0, 0, 0, 0.08);
    }
    
    .timer-warning {
      color: #b31b1b;
      font-weight: 700;
      animation: pulse 1s infinite;
    }
    
    @keyframes pulse {
      0%, 100% { opacity: 1; }
      50% { opacity: 0.7; }
    }
  "
)
```

## Launching the Study

### 1. Basic Launch

```{r basic_launch}
# Launch the study with the monochrome theme
launch_study(
  config = study_config,
  item_bank = r_testing_items,
  port = 3838,
  launch_browser = TRUE
)
```

### 2. Advanced Launch with Custom Features

```{r advanced_launch}
# Launch with additional features
launch_study(
  config = enhanced_config,
  item_bank = r_testing_items,
  
  # Server Configuration
  port = 3838,
  host = "0.0.0.0",
  launch_browser = TRUE,
  
  # Data Collection
  data_collection = list(
    save_responses = TRUE,
    save_path = "study_responses/",
    save_format = "csv",
    include_timestamps = TRUE,
    include_session_info = TRUE
  ),
  
  # Quality Control
  quality_control = list(
    detect_rapid_responses = TRUE,
    min_response_time = 2,  # seconds
    detect_straight_lining = TRUE,
    require_progress_minimum = 0.8
  ),
  
  # Accessibility Features
  accessibility = list(
    enable_screen_reader = TRUE,
    enable_keyboard_navigation = TRUE,
    high_contrast_mode = TRUE,
    font_size_adjustment = TRUE
  )
)
```

## Key Features Demonstrated

### 1. Monochrome Theme Benefits

The Monochrome theme provides:

- **Professional Appearance**: Clean black and white design suitable for academic research
- **High Contrast**: Excellent readability and accessibility compliance
- **Elegant Typography**: Lora serif headings with Inter sans-serif body text
- **Sophisticated Animations**: Subtle hover effects and transitions
- **Responsive Design**: Optimal display across devices
- **Print-Friendly**: Clean printing with appropriate styles

### 2. Psychological Assessment Features

- **Likert Scale Support**: Proper handling of ordinal response data
- **Timer Integration**: Built-in 20-minute time limit with warnings
- **Progress Tracking**: Visual progress indicators with animations
- **Skip Options**: Flexible response requirements
- **Demographic Collection**: Comprehensive participant profiling
- **Results Export**: Downloadable response data

### 3. Research Compliance

- **Informed Consent**: Built-in consent form handling
- **Data Anonymization**: Automatic anonymization of responses
- **Session Management**: Secure session handling and storage
- **Accessibility**: WCAG 2.1 AA compliance
- **Quality Control**: Response validation and quality checks

## Customization Options

### 1. Theme Customization

```{r theme_customization}
# Customize the Monochrome theme
custom_monochrome_config <- create_study_config(
  name = "Custom Monochrome Study",
  theme = "Monochrome",
  
  # Custom theme modifications
  theme_config = list(
    primary_color = "#2c2c2c",
    accent_color = "#888888",
    font_family_heading = "Georgia, serif",
    font_family_body = "Arial, sans-serif",
    container_width = "900px",
    card_padding = "3rem"
  ),
  
  # Additional customizations...
  demographics = c("age_range", "r_experience"),
  model = "GRM",
  adaptive = FALSE
)
```

### 2. Question Customization

```{r question_customization}
# Create custom question types
custom_items <- data.frame(
  item_id = 1:8,
  item_text = c(
    # Original questions
    "I feel confident in my ability to write effective unit tests using testthat.",
    "Using devtools to manage and test R packages is intuitive and efficient.",
    "Interpreting code coverage reports from covr helps me improve my testing strategy.",
    "Testing R packages reduces my stress when preparing for a package release.",
    "I feel motivated to incorporate regular testing into my R package development process.",
    "The documentation for testthat is clear and supports my testing needs.",
    
    # Additional questions
    "I find R package testing workflows to be time-consuming and tedious.",
    "I would recommend R testing tools to other developers in my field."
  ),
  
  # Add reverse-coded item (item 7)
  reverse_coded = c(rep(FALSE, 6), TRUE, FALSE),
  
  # Question categories
  category = c(
    "Confidence", "Usability", "Utility", "Stress", "Motivation", "Documentation",
    "Efficiency", "Recommendation"
  ),
  
  # IRT parameters
  a = rep(1.5, 8),
  b1 = rep(-2, 8),
  b2 = rep(-1, 8),
  b3 = rep(0, 8),
  b4 = rep(1, 8),
  
  response_options = I(rep(list(c(
    "Strongly Disagree", "Disagree", "Neutral", "Agree", "Strongly Agree"
  )), 8))
)
```

### 3. Results Customization

```{r results_customization}
# Custom results processing
custom_results_config <- create_study_config(
  name = "R Testing Psychology Study",
  theme = "Monochrome",
  
  # Custom results function
  results_processing_fun = function(responses, demographics, session_data) {
    # Calculate subscale scores
    confidence_items <- c(1, 2, 3)
    motivation_items <- c(4, 5, 8)
    usability_items <- c(2, 6, 7)
    
    # Reverse code item 7
    responses[7] <- 6 - responses[7]  # Reverse 1-5 scale
    
    # Calculate subscale means
    confidence_score <- mean(responses[confidence_items], na.rm = TRUE)
    motivation_score <- mean(responses[motivation_items], na.rm = TRUE)
    usability_score <- mean(responses[usability_items], na.rm = TRUE)
    
    # Create results summary
    results_summary <- list(
      total_score = mean(responses, na.rm = TRUE),
      confidence_score = confidence_score,
      motivation_score = motivation_score,
      usability_score = usability_score,
      completion_rate = sum(!is.na(responses)) / length(responses),
      session_duration = session_data$duration,
      demographic_profile = demographics
    )
    
    return(results_summary)
  },
  
  # Other configurations...
  demographics = c("age_range", "r_experience"),
  model = "GRM",
  adaptive = FALSE
)
```

## Complete Example

Here's a complete, ready-to-run example:

```{r complete_example}
library(inrep)

# 1. Create item bank
r_testing_items <- data.frame(
  item_id = 1:6,
  item_text = c(
    "I feel confident in my ability to write effective unit tests using testthat.",
    "Using devtools to manage and test R packages is intuitive and efficient.",
    "Interpreting code coverage reports from covr helps me improve my testing strategy.",
    "Testing R packages reduces my stress when preparing for a package release.",
    "I feel motivated to incorporate regular testing into my R package development process.",
    "The documentation for testthat is clear and supports my testing needs."
  ),
  a = rep(1.5, 6),
  b1 = rep(-2, 6),
  b2 = rep(-1, 6),
  b3 = rep(0, 6),
  b4 = rep(1, 6),
  response_options = I(rep(list(c(
    "Strongly Disagree", "Disagree", "Neutral", "Agree", "Strongly Agree"
  )), 6))
)

# 2. Create study configuration
study_config <- create_study_config(
  name = "Psychological Study on R Package Testing Experience",
  theme = "Monochrome",
  demographics = c("age_range", "r_experience"),
  input_types = list(age_range = "radio", r_experience = "radio"),
  study_key = "r_testing_psychology_2025",
  model = "GRM",
  adaptive = FALSE,
  max_items = 6,
  min_items = 6,
  session_timeout = 20,
  show_timer = TRUE,
  demographic_configs = list(
    age_range = list(
      question = "What is your age range?",
      options = c("18-24", "25-34", "35-44", "45-54", "55+")
    ),
    r_experience = list(
      question = "How many years have you been using R for package development?",
      options = c("Less than 1 year", "1-3 years", "3-5 years", "5+ years")
    )
  )
)

# 3. Launch the study
launch_study(
  config = study_config,
  item_bank = r_testing_items,
  port = 3838,
  launch_browser = TRUE
)
```

## Conclusion

This example demonstrates how the `inrep` package with the **Monochrome** theme can be used to create sophisticated psychological studies. The combination of:

- **Elegant Design**: Professional monochrome aesthetic
- **Research Features**: Timer, progress tracking, demographics collection
- **Accessibility**: WCAG 2.1 AA compliance
- **Flexibility**: Customizable themes and configurations
- **Data Quality**: Built-in validation and quality control

Makes it an ideal platform for conducting psychological research on software development practices and developer experiences.

The **Monochrome** theme specifically provides the professional, distraction-free environment that psychological studies require, while maintaining excellent usability and accessibility standards.
