---
title: "Quick Start Guide: inrep Package"
author: "inrep Development Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_width: 10
    fig_height = 6
    df_print: paged
vignette: >
  %\VignetteIndexEntry{Quick Start Guide: inrep Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  eval = TRUE,
  fig.width = 10,
  fig.height = 6,
  fig.align = "center",
  warning = FALSE,
  message = FALSE
)

# Load required packages
library(inrep)
library(ggplot2)
library(dplyr)

# Set seed for reproducibility
set.seed(123)
```

# Quick Start Guide

This guide gets you up and running with the `inrep` package in minutes. Whether you're conducting educational assessments, psychological research, or clinical evaluations, this guide will help you create your first adaptive test quickly and efficiently.

## Installation

```{r eval=FALSE}
# Install from CRAN (when available)
install.packages("inrep")

# Or install from local source
devtools::install_local("path/to/inrep")

# Or install development version from GitHub
devtools::install_github("your-repo/inrep")

# Load the package
library(inrep)
```

## System Requirements

Before starting, ensure you have:
- R version 4.0.0 or higher
- Required packages: TAM, shiny, ggplot2, dplyr
- For parallel processing: future, future.apply
- For advanced features: mirt (optional)

## 5-Minute Setup

### Step 1: Load Data

```{r}
# Load mathematics items
data(math_items)

# Or load personality items
# bfi_items is automatically available via LazyData
```

### Step 2: Create Configuration

```{r}
# Simple math test
config <- create_study_config(
  study_name = "Quick Math Test",
  item_bank = "math_items",
  max_items = 10,
  min_items = 5
)
```

### Step 3: Launch Assessment

```{r}
# Start the assessment
launch_study(config)
```

## Common Use Cases

### Educational Testing

```{r}
# Create educational math assessment
edu_config <- create_study_config(
  study_name = "Grade 8 Math Assessment",
  item_bank = "math_items",
  max_items = 15,
  adaptive_selection = TRUE,
  demographics = list(
    grade = list(type = "select", options = c("8")),
    student_id = list(type = "text", required = TRUE)
  )
)

launch_study(edu_config)
```

### Personality Research

```{r}
# Create personality study
personality_config <- create_study_config(
  study_name = "Personality Research",
  item_bank = "bfi_items",
  max_items = 44,
  demographics = list(
    age = list(type = "numeric", required = TRUE),
    consent = list(type = "checkbox", required = TRUE)
  )
)

launch_study(personality_config)
```

### Adaptive Testing

```{r}
# Create adaptive test
adaptive_config <- create_study_config(
  study_name = "Adaptive Math Test",
  item_bank = "math_items",
  adaptive_selection = TRUE,
  selection_method = "maximum_info",
  stopping_criterion = "se_threshold",
  se_threshold = 0.4,
  max_items = 20,
  min_items = 8
)

launch_study(adaptive_config)
```

## Essential Functions

### Configuration
- `create_study_config()`: Main configuration function
- `validate_config()`: Validate configuration settings

### Assessment
- `launch_study()`: Start assessment interface
- `estimate_ability()`: Estimate participant ability

### Analysis
- `analyze_responses()`: Analyze response patterns
- `generate_report()`: Create assessment reports

## Next Steps

1. **Read the comprehensive guide**: See `vignette("comprehensive_adaptive_testing")`
2. **Compare domains**: See `vignette("mathematics_vs_personality")`
3. **Customize your assessment**: Explore advanced configuration options
4. **Analyze results**: Use built-in analysis functions

## Getting Help

- Documentation: `?inrep` or `help(package = "inrep")`
- Examples: `example("create_study_config")`
- Vignettes: `vignette(package = "inrep")`

## Common Issues

**Q: Assessment won't launch**
A: Check that all required columns are present in your item bank

**Q: TAM errors**
A: Ensure the TAM package is installed and loaded

**Q: Configuration errors**
A: Use `validate_config()` to check your settings

Ready to start? Try the examples above!
